---
title: "Business Conditions Survey"
subtitle: COVID response
output:
  html_document:
    df_print: paged
---


```{r setup}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)

library(cansim)
```

## Business Conditions Survey: COVID


[The Daily: Canadian Survey on Business Conditions: Impact of COVID-19 on businesses in Canada, May 2020](https://www150.statcan.gc.ca/n1/daily-quotidien/200714/dq200714a-eng.htm)

Meta details
https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5318



### Data tables

[comprehensive list of pandemic-related business conditions survey tables](https://www150.statcan.gc.ca/n1/daily-quotidien/200714/dq200714a-cansim-eng.htm)

Personal protective equipment or supply needs, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0259-01

Geography: Canada
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310025901


Difficulty procuring personal protective equipment or supplies, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0260-01

Geography: Canada
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310026001



Sources of personal protective equipment or supplies, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0261-01

Geography: Canada



```{r}

df_3310025901 <- cansim::get_cansim("33-10-0259-01")

df_3310025901 <-
df_3310025901 %>% 
  janitor::clean_names()


df_3310026001 <- cansim::get_cansim("33-10-0260-01")

df_3310026001 <-
df_3310026001 %>% 
  janitor::clean_names()


df_3310026101 <- cansim::get_cansim("33-10-0261-01")

df_3310026101 <-
df_3310026101 %>% 
  janitor::clean_names()

```


### personal_protective_equipment_or_supplies_businesses_need_or_expect_to_need
 

British Columbia

```{r}

df_3310025901 %>%
  filter(business_characteristics == "British Columbia") %>% 
  select(personal_protective_equipment_or_supplies_businesses_need_or_expect_to_need,
         value)

```

## Businesses experiencing or expecting to experience difficulty procuring protective equipment or supplies

British Columbia

```{r}

df_3310026001 %>%
  filter(business_characteristics == "British Columbia") %>% 
  select(businesses_experiencing_or_expecting_to_experience_difficulty_procuring_protective_equipment_or_supplies)

```

British Columbia

(also by a variety of other characteristics, including industy sector (2 digit NAICS), number of employees, ownership characteristic, exporter profile--but no crosstabs)

```{r}

unique(df_3310026001$business_characteristics)

df_3310026001 %>%
  filter(business_characteristics == "British Columbia") %>% 
  select(businesses_experiencing_or_expecting_to_experience_difficulty_procuring_protective_equipment_or_supplies, value)

```

## source_s_from_which_businesses_have_obtained_or_expect_to_obtain_personal_protective_equipment

```{r}

df_3310026101 %>%
  filter(business_characteristics == "British Columbia") %>% 
  select(source_s_from_which_businesses_have_obtained_or_expect_to_obtain_personal_protective_equipment_or_supplies, value)

```
