---
title: "Business Conditions Survey"
subtitle: "COVID response"
output: html_notebook
---


```{r setup}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)

library(cansim)
```

## Business Conditions Survey: COVID

### Data tables

[comprehensive list of pandemic-related business conditions survey tables](https://www150.statcan.gc.ca/n1/daily-quotidien/200714/dq200714a-cansim-eng.htm)

Personal protective equipment or supply needs, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0259-01

Geography: Canada
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310025901


Difficulty procuring personal protective equipment or supplies, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0260-01

Geography: Canada
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310026001



Sources of personal protective equipment or supplies, by business characteristics1 2
Frequency: Occasional Monthly

Table: 33-10-0261-01

Geography: Canada



```{r}

df_3310025901 <- cansim::get_cansim("33-10-0259-01")

df_3310025901 <-
df_3310025901 %>% 
  janitor::clean_names()


df_3310026001 <- cansim::get_cansim("33-10-0260-01")

df_3310026001 <-
df_3310026001 %>% 
  janitor::clean_names()


df_3310026101 <- cansim::get_cansim("33-10-0261-01")

df_3310026101 <-
df_3310026101 %>% 
  janitor::clean_names()

```
