---
title: "LFS - occupation summary"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)

library(cansim)
```


## Data sources


Labour force characteristics by occupation, monthly, unadjusted for seasonality (x 1,000)
Frequency: Monthly

Table: 14-10-0296-01 (formerly CANSIM 282-0141)

Geography: Canada, Province or territory



```{r}

df_1410029601 <- get_cansim("14-10-0296-01")

df_1410029601 <- janitor::clean_names(df_1410029601)

df_1410029601 <- df_1410029601 %>% 
  mutate(ref_date_2 = glue::glue("{ref_date}-01")) %>% 
  mutate(ref_date_2 = ymd(ref_date_2)) 

write_rds(df_1410029601, here::here("data", "df_1410029601.rds"))

```


```{r}
df_14100194 <- janitor::clean_names(df_14100194)

df_14100194 <- df_14100194 %>% 
  mutate(ref_date_2 = glue::glue("{ref_date}-01-01")) %>% 
  mutate(ref_date_2 = ymd(ref_date_2)) %>% 
  mutate(presence_of_children = as.factor(presence_of_children))


head(df_14100194)

unique(df_14100194$presence_of_children)

df_14100194$presence_of_children <- 
  fct_relevel(df_14100194$presence_of_children, 
            levels = c("Preschoolers, under 5 years",
                       "Children 5 to 12 years",
                       "Children 13 years and over"))

levels(df_14100194$presence_of_children)

```



Actual hours worked by occupation, annual
Frequency: Annual

Table: 14-10-0301-01 (formerly CANSIM 282-0146)

Geography: Canada, Province or territory


Labour force characteristics by occupation, annual
Frequency: Annual

Table: 14-10-0335-01

Geography: Canada, Province or territory


