<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->

---
title: "construction investment"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

# Investment in Building Construction


This set of scripts creates summary tables and graphs plotting the monthly SEPH data collected and reported by Statistics Canada.

#### packages

```{r setup, packages}

# tidyverse
library(tidyverse)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library(scales)

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

# utilities
library(brickr)
library(here)
library(janitor)

### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)

```

---

## Background


The Daily, 2019-07-25: Investment in building construction, June 2019
https://www150.statcan.gc.ca/n1/daily-quotidien/190821/dq190821b-eng.htm?CMP=mstatcan



## -- Data

* Investment in Building Construction

  - Table: 34-10-0175-01
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3410017501


### Read the data


**NOTE:**

This script uses the [R package {cansim}](https://github.com/mountainMath/cansim)


```{r data_read}

data_34_10_0175 <- get_cansim("34-10-0175-01")

# {cansim} giving me grief today...time to investigate and perhaps file an issue
# data downloaded from https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3410017501 
# (see "Download options" button on right)
# unzipped files saved in project folder, in sub-folder \data\

#data_34_10_0175 <- read_csv(here("data", "34100175.csv"))

data_coninv <- data_34_10_0175 %>%
  janitor::clean_names() %>%
  mutate(ref_date = ymd(ref_date, truncated = 2))

ls.str(data_coninv)


```


### filters & groups

* geo == `British Columbia`

* investment_value

* type_of_work

* type_of_structure

```{r}

data_coninv_bc <- data_coninv %>%
  filter(geo == "British Columbia")

unique(data_coninv$investment_value)

unique(data_coninv$type_of_work)

unique(data_coninv$type_of_structure)

last_date <- max(data_coninv_bc$ref_date)

last_date

```


```{r}

data_coninv_bc %>% 
  filter(ref_date == last_date)

```



```{r}

data_coninv_bc %>%
  filter(type_of_work == "Types of work, total",
         investment_value == "Seasonally adjusted - current",
         type_of_structure %in% c("Total residential and non-residential",
                                  "Total residential",
                                  "Single dwelling building total",
                                  "Multiple dwelling building total",
                                  "Total non-residential",
                                  "Total industrial",
                                  "Total commercial",
                                  "Total institutional and governmental")) %>% 
  group_by(type_of_structure) %>% 
  mutate(MOM = value - lag(value, 1),
         YOY = value - lag(value, 12)) %>% 
  mutate(MOM_pct = (MOM/value * 100),
         YOY_pct = (YOY/value * 100)) %>% 
  ungroup() %>% 
  select(ref_date, geo, type_of_structure,
         value, MOM, MOM_pct, YOY, YOY_pct) %>% 
  filter(ref_date == last_date)

```




```{r}


data_coninv_bc <- data_coninv_bc %>%
  filter(ref_date == last_date,
         investment_value == "Unadjusted - current",
         type_of_work == "Types of work, total")

data_coninv_bc

unique(data_coninv_bc$type_of_structure)

data_summary <-
data_coninv_bc %>%
  mutate(type_grp = case_when(
    type_of_structure == "Single dwelling building total" ~ "Residential",
    type_of_structure == "Multiple dwelling building total" ~ "Residential",
    type_of_structure == "Total industrial" ~ "Non-residential",
    type_of_structure == "Total commercial" ~ "Non-residential",
    type_of_structure == "Total institutional and governmental" ~ "Non-residential",
    TRUE ~ NA_character_
  )) %>%
  mutate(type_grp = as_factor(type_grp),
         type_of_structure = as_factor(type_of_structure))

data_summary$type_of_structure <-
data_summary$type_of_structure %>%
  fct_recode(
    "Single dwelling building" = "Single dwelling building total",
    "Multiple dwelling building" = "Multiple dwelling building total",
    "Industrial" = "Total industrial",
    "Commercial" = "Total commercial",
    "Institutional & governmental" = "Total institutional and governmental"
  )

data_chart <- data_summary %>%
  filter(!is.na(type_grp)) %>%
  group_by(type_grp, type_of_structure) %>%
  summarise(valuesum = sum(value, na.rm = TRUE)) %>%
  arrange(valuesum)

data_chart

```

## plot

```{r}

coninv_plot <- ggplot(data_chart) +
  aes(x=type_of_structure, y=valuesum) + 
  geom_brick_col(two_knob = FALSE) +
  coord_brick_flip() 
#  scale_fill_brick() 
#  theme_brick()

coninv_plot

```

```{r}

coninv_plot <- ggplot(data_chart) +
  aes(x=type_of_structure, y=valuesum) + 
  geom_brick_col(two_knob = FALSE) +
  coord_brick_flip() +
  facet_grid(rows = vars(type_grp))
#  scale_fill_brick() 
#  theme_brick()

coninv_plot

```



```{r}

coninv_plot <- ggplot(data_chart) +
  aes(x=reorder(type_of_structure, valuesum), y=valuesum) + 
  geom_brick_col(aes(fill = type_grp), two_knob = FALSE) +
  scale_fill_brick() +
  coord_brick_flip() 
#  theme_brick()

coninv_plot

```

plot, with formatting

```{r}

coninv_plot +
  labs(title = "Investment in Building Construction",
       subtitle = "British Columbia, November 2019",
       caption = "Source: Statistics Canada. Table 34-10-0175-01; Investment in Building Construction") +
  xlab(element_blank()) +
  ylab("$ (unadjusted, current)") +
  scale_y_continuous(labels = comma) +
  # theme
  theme(legend.title = element_blank(),
        legend.position = "bottom")

```




```{r}

ggsave(filename = "construction_investment_plot.png", 
       width = 8, height = 6)



```


-30-