---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Employment Insurance


This set of scripts creates summary tables and graphs plotting the Monthly Survey of Manufacturing data collected and reported by Statistics Canada.

#### packages

```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

```

---

## Background

The Daily, 2018-09-20

* https://www150.statcan.gc.ca/n1/daily-quotidien/180920/dq180920a-eng.htm?HPA=1 



## data

There are 12 data tables associated with this:
https://www150.statcan.gc.ca/n1/daily-quotidien/180920/dq180920a-cansim-eng.htm


Our focus is on the seasonally adjusted series.



Provincial:

* Employment insurance claims received by province and territory, monthly, seasonally adjusted

  - Table: 14-10-0005-01 (formerly CANSIM  276-0004)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410000501 

* Employment insurance regular income benefit characteristics, monthly, seasonally adjusted

  - Table: 14-10-0008-01 (formerly CANSIM  276-0018)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410000801
  
* Employment insurance beneficiaries (regular benefits) by province and territory, monthly, seasonally adjusted

  - Table: 14-10-0011-01 (formerly CANSIM  276-0022)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410001101
  
* Employment insurance beneficiaries (regular benefits) by census metropolitan category[^1], monthly, seasonally adjusted

  - Table: 14-10-0013-01 (formerly CANSIM  276-0034)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410001301
  
  
* Employment insurance beneficiaries (regular benefits) by province, territory and occupation, monthly, seasonally adjusted

  - Table: 14-10-0337-01 (formerly CANSIM  276-0043)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410033701
  


###

Read the data


```{r}

data_10_0005 <- get_cansim("14-10-0005-01")

data_10_0005 <- data_10_0005 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 
  
data_10_0008 <- get_cansim("14-10-0008-01")

data_10_0008 <- data_10_0008 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

data_10_0011 <- get_cansim("14-10-0011-01")

data_10_0011 <- data_10_0011 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

data_10_0013 <- get_cansim("14-10-0013-01")

data_10_0013 <- data_10_0013 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

data_10_0337 <- get_cansim("14-10-0337-01")

data_10_00337 <- data_10_0337 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

```



this month

```{r}

filenamenum <- "data_10_0011"

print(glue("TABLE:  ", filenamenum))

ls.str(data_10_0011)

unique(data_10_0011$`Beneficiary detail`)


data_10_0011 %>%
  filter(GEO %in% c("Canada", "British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over",
    REF_DATE == "2018-07-01")



data_10_0011 %>%
  filter(GEO %in% c("Canada", "British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(REF_DATE)) %>%
  filter(REF_DATE == "2018-07-01")

```


plot: BC trend

```{r}

data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  select(REF_DATE, VALUE, MOM_pct, MOM_val) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2006-01-01")


data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5) 


```


```{r}

unique(year(data_10_0011$REF_DATE))

data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)

```


```{r}

dataplot <- data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5, colour="#325A80") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)
  



dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(30000, 100000)) +
#  scale_colour_manual(name=NULL,
#                      breaks=c("Canada", "British Columbia"),
#                      labels=c("Canada", "British Columbia"), 
#                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=10)
  )
#
dataplot2 

# experiments with ggplot2's new subtitle and caption options

EI_title <- as.character("Employment Insurance beneficiaries (seasonally adjusted), B.C.")
EI_subtitle <- as.character("January 2010 = 100")
EI_caption <- as.character("Source: Statistics Canada, CANSIM table 14-10-0011-01")
EI_yaxis <- as.character("Number of EI beneficiaries")

# add titles / X-Y axis labels / caption
EI_plot <- dataplot2 +
  labs(title = EI_title,
#       subtitle = EI_subtitle,
       caption = EI_caption, 
       x = NULL, y = EI_yaxis) 

EI_plot

ggsave(filename = "EI_plot.png", plot = EI_plot,
       width = 8, height = 6)

```


### B.C. Geography


```{r}


unique(data_10_0013$GeoUID)

CMA_CA_list <- c("[11124]",
                 "[59]", "[59A]", "[932]", "[915]", "[933]", "[935]", 
                 "[59B]", "[944]", "[930]", "[943]", "[905]", 
                 "[975]", "[937]", "[977]", "[925]", "[938]", "[939]", 
                 "[913]", "[940]", "[945]", "[970]", "[955]", "[952]", 
                 "[920]", "[934]", "[965]", "[918]", "[950]", 
                 "[59C]", "[59996]", "[59997]", "[59998]", "[59999]")


filenamenum <- "data_10_0013"

print(glue("TABLE:  ", filenamenum))

ls.str(data_10_0013)



data_10_0013 %>%
  filter(GeoUID %in% CMA_CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#  arrange(desc(REF_DATE)) %>%
  filter(REF_DATE == "2018-07-01")



data_10_0013 %>%
  filter(GeoUID %in% CMA_CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_val)) %>%
  filter(REF_DATE == "2018-07-01")

CMA_list <- c("[59A]", "[932]", "[915]", "[933]", "[935]")

data_10_0013 %>%
  filter(GeoUID %in% CMA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_pct)) %>%
  filter(REF_DATE == "2018-07-01")


CA_list <- c("[59B]", "[944]", "[930]", "[943]", "[905]", 
             "[975]", "[937]", "[977]", "[925]", "[938]", "[939]", 
             "[913]", "[940]", "[945]", "[970]", "[955]", "[952]", 
             "[920]", "[934]", "[965]", "[918]", "[950]")


data_10_0013 %>%
  filter(GeoUID %in% CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_pct)) %>%
  filter(REF_DATE == "2018-07-01")

```



## July 2018 text

Issue 18-nnn Employment Insurance

September 20, 2018

The number of people receiving employment insurance (EI) benefits in British Columbia rose 2.2% in July compared to June (seasonally adjusted), up by 900 to reach 42,570 individuals. This was the first increase seen since November 2017, and only the third in the 21 months since November 2016. In October 2016, the total number of people in B.C. receiving EI benefits was 57,000. Looking back over the last decade, the number of EI recipients in B.C. peaked at nearly 98,000 in June 2009.

The increase in beneficiaries in British Columbia was observed in all census metropolitan areas in B.C. (+3.1%, or 680 persons). The largest increase was seen in Victoria, which added 290 persons (+14.5%) and Abbotsford (+140 persons, +7.7%). Census agglomerations also saw an aggregate increase (370 persons or +3.4%), although changes in the number of EI beneficiaries ranged from an increase of 31% (+110 persons) in Duncan to drop of 10.5% (-20 persons) in Dawson Creek. Areas outside census metropolitan areas and census agglomerations saw a drop of 140 persons or -1.5%.

Nationally, the number of people receiving EI benefits rose 2.4% (an increase of 11,000 people) compared to June. The three provinces with rates of increase higher than the national rate were Manitoba (+18.2%), Quebec (+4.5%), and Prince Edward Island (+4.0%). Statistics Canada reports that the number of EI beneficiaries "fell slightly in Saskatchewan (-1.3%) and Newfoundland and Labrador (-1.0%), while it was little changed in Nova Scotia and Ontario."


Data Source: Statistics Canada (https://www150.statcan.gc.ca/n1/daily-quotidien/180920/dq180920a-eng.htm)




***

### footnotes

[^1]
>Geographical definitions: A census metropolitan area (CMA) or a census agglomeration (CA) is formed by one or more adjacent municipalities centred on a population centre. A CMA must have a total population of at least 100,000. A CA must have a population of at least 10,000. See Standard Geographical Classification 2011 – definitions for more information.
