---
title: "Residency status of residential properties"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Residential status of residential properties


This set of scripts creates summary tables and graphs plotting Canadian Housing Statistics Program (CHSP) data collected and reported by Statistics Canada.


#### packages

```{r packages}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)


### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)

```

---

## Background



## data

The table used:

* Residency status of residential properties by property type and period of construction, provinces of Nova Scotia, Ontario and British Columbia

  - Table: 46-10-0018-01 
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=4610001801



### Read the data


**NOTE:**

This script uses the [R package `cansim`](https://github.com/mountainMath/cansim)


```{r data_read}

data_46_10_0018_01 <- get_cansim("46-10-0018-01")


housing_data <- data_46_10_0018_01

ls.str(housing_data)

```


---

#### some quick summaries of the data 





-30-