---
title: "Residency status of residential properties"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Residential status of residential properties


This set of scripts creates summary tables and graphs plotting Canadian Housing Statistics Program (CHSP) data collected and reported by Statistics Canada.


#### packages

```{r packages}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library(scales)

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)


### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)

```

---

## Background



## data

The table used:

* Residency status of residential properties by property type and period of construction, provinces of Nova Scotia, Ontario and British Columbia

  - Table: 46-10-0018-01 
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=4610001801


Geography references

* [Census Geography](https://www12.statcan.gc.ca/census-recensement/2016/geo/index-eng.cfm)

* [Geography Catalogue, Census year 2016](https://www150.statcan.gc.ca/n1/pub/92-196-x/92-196-x2016001-eng.htm)

* [Correspondence Files, Reference Guide](https://www150.statcan.gc.ca/n1/pub/92-156-g/92-156-g2016001-eng.htm)




### Read the data


**NOTE:**

This script uses the [R package `cansim`](https://github.com/mountainMath/cansim)


```{r data_read}

CHSP_data_CDN <- get_cansim("46-10-0018-01")


```


```{r}

CHSP_data_CDN %>%
  select(GEO, GeoUID) %>%
  group_by(GEO, GeoUID) %>%
  tally() %>%
  arrange(GeoUID)


# create level of geography variable

CMA_list <- CHSP_data_CDN %>%
  filter(str_detect(GeoUID, "\\[\\d{3}\\]")) %>%
  pull(DGUID) %>%
  unique()


CHSP_data_CDN <- 
CHSP_data_CDN %>%
  mutate(GEO_level =
           case_when(
             DGUID %in% CMA_list ~ "CMA",
             is.na(DGUID) ~ "CMA",
             str_detect(GeoUID, "\\[\\d\\d\\]") ~ "PROV",
             TRUE ~ "CSD"
             )
           )


# quick test
CHSP_data_CDN %>%
  filter(GEO_level == "CMA") %>%
  pull(GEO) %>%
  unique()


# --------
# create GEO_prov


CHSP_data_CDN <-
  CHSP_data_CDN %>%
  mutate(GEO_prov =
           str_sub(GeoUID, start = 2, end = 3)
         )

CHSP_data_CDN %>%
  group_by(GEO_prov) %>%
  tally()

# no, that's not quite right--the CMAs are messing it up

CMA_names_list <- CHSP_data_CDN %>%
  filter(GEO_level == "CMA") %>%
  pull(GEO) %>%
  unique()

CMA_names_list

CHSP_data_CDN <-
CHSP_data_CDN %>%
  mutate(GEO_prov =
           case_when(
             GEO == "Halifax, census metropolitan area" ~ "12",                       
             GEO == "Nova Scotia, outside of census metropolitan areas" ~ "12",       
             GEO == "Barrie, census metropolitan area" ~ "35",                        
             GEO == "Belleville, census metropolitan area" ~ "35",                    
             GEO == "Brantford, census metropolitan area" ~ "35",                     
             GEO == "Greater Sudbury, census metropolitan area" ~ "35",               
             GEO == "Guelph, census metropolitan area" ~ "35",                        
             GEO == "Hamilton, census metropolitan area" ~ "35",                      
             GEO == "Kingston, census metropolitan area" ~ "35",                      
             GEO == "Kitchener-Cambridge-Waterloo, census metropolitan area" ~ "35",  
             GEO == "London, census metropolitan area" ~ "35",                        
             GEO == "Oshawa, census metropolitan area" ~ "35",                        
             GEO == "Ottawa-Gatineau (Ontario part), census metropolitan area" ~ "35",
             GEO == "Peterborough, census metropolitan area" ~ "35",                  
             GEO == "St. Catharines-Niagara, census metropolitan area" ~ "35",        
             GEO == "Thunder Bay, census metropolitan area" ~ "35",                   
             GEO == "Toronto, census metropolitan area" ~ "35",                       
             GEO == "Windsor, census metropolitan area" ~ "35",                       
             GEO == "Ontario, outside of census metropolitan areas" ~ "35",           
             GEO == "Abbotsford-Mission, census metropolitan area" ~ "59",            
             GEO == "Kelowna, census metropolitan area" ~ "59",                       
             GEO == "Vancouver, census metropolitan area" ~ "59",                     
             GEO == "Victoria, census metropolitan area" ~ "59",                      
             GEO == "British Columbia, outside of census metropolitan areas" ~ "59",
             TRUE ~ str_sub(GeoUID, start = 2, end = 3)
           ))


CHSP_data_CDN %>%
  group_by(GEO_prov) %>%
  tally()

```



---

#### some quick summaries of the data 

```{r}

housing_data_BCprov <- CHSP_data_CDN %>%
  filter(GEO == "British Columbia") 

housing_data_BC_all <- 
CHSP_data_CDN %>%
  filter(!str_detect(GeoUID, "\\[595\\]"))%>%
  filter(str_detect(GeoUID, "\\[59")) 

```



```{r}

unique(housing_data_BC_all$GEO)

unique(housing_data_BC_all$GeoUID)

```


```{r}
housing_data_BC %>%
  filter(Estimates == "Median assessment value",
         `Residency status` != "Total, all residency status categories") %>%
  select(GEO, `Property type`, `Period of construction`, `Residency status`,
         Estimates, UOM, VALUE) 



```




-30-