---
title: "International merchandise trade - seasonal adjustment"
output:
  html_document:
    df_print: paged
---
<!--
Copyright 2018-2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


This set of scripts 

* runs the seaonal adjustment algorithm

* creates summary tables and graphs 

plotting International Merchandise Trade data for Canada and British Columbia collected and reported by Statistics Canada.


#### Reference material

See https://bookdown.org/martin_monkman/DataScienceResources_book/quantitativemethods2.html#seasonal-adjustment 


#### packages



```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library("ggplot2") # NOTE: requires ggplot 2.2.0 
#library("lubridate")

# extending ggplot2
library(scales)

# monthly data series
#library(zoo)

#
library(cansim)

# seasonal adjustment
library(seasonal)

```

---

### data


ARCHIVED DATA SERIES TO 2016-12 
read the data
- file 03270046-eng.csv download of CANSIM table 327-0046
- http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=3270046&&pattern=&stByVal=1&p1=1&p2=1&tabMode=dataTable&csid=


#### download and save

**data source**

International merchandise trade by province, commodity, and Principal Trading Partners (x 1,000)

Frequency: Monthly

Table: 12-10-0119-01

Geography: Canada, Province or territory



https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1210011901




```{r}

table_id = "12-10-0119-01"


thedata <- get_cansim(table_id)
thedata <- thedata %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

get_cansim_table_overview(table_id)


thedata


```




#### data wrangling


geography list

* note: need stringr to allow filtering of `GeoUID`by 2-digit code (e.g., BC = 59)

```{r}

unique(thedata$GEO)
unique(thedata$Trade)


thedata %>%
#  filter(GeoUID == ) %>%
  group_by(GEO, GeoUID) %>%
  tally()

```





```{r}

#startdate <- as.Date("2007-01-01")


# filter to have BC and Canada
thedata_CANBC_exp <- thedata %>%
#  filter(REF_DATE >= startdate) %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `Trade` == "Domestic export",
         `Principal trading partners` %in% c("All countries", "United States"))

thedata_CANBC_exp

CC_list <- c("[C11]", "[C12]", "[C13]", "[C14]",
             "[C15]", "[C16]", "[C17]", "[C18]",
             "[C19]", "[C21]", "[C22]", "[C23]")


thedata_components <- thedata_CANBC_exp %>%
  filter(`Classification Code for North American Product Classification System (NAPCS)` %in% CC_list)

unique(thedata_components$`Classification Code for North American Product Classification System (NAPCS)`)

thedata_components_singlemonth <- thedata_components %>%
  filter(REF_DATE == "2019-04-01")


thedata_components_singlemonth

```


#### data check

Do the components sum to the total?

```{r}

# components

thedata_components_singlemonth %>%
  group_by(REF_DATE, GEO, `Principal trading partners`) %>%
  summarise(month_total = sum(VALUE))


# total

thedata %>%
#  filter(REF_DATE >= startdate) %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `Trade` == "Domestic export",
         `Principal trading partners` %in% c("All countries", "United States"),
         REF_DATE == "2019-04-01",
         `North American Product Classification System (NAPCS)` == "Total of all merchandise") %>%
  select(REF_DATE, GEO, `Principal trading partners`, VALUE)


```

(check B.C. against https://www.bcstats.gov.bc.ca/Files/049a978c-7975-40db-9463-31ce7fde2275/ExportsData1512.xls -- works!)


## Seasonal adjustment

Need to define values as time series


