---
title: "Labour Productivity"
output:
  html_document:
    df_print: paged
---
<!--
Copyright 2018-2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


This set of scripts creates summary tables and graphs plotting Labour Productivity data collected and reported by Statistics Canada.

This uses the `cansim` package to pull the data from CANSIM.


#### packages



```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library("ggplot2") # NOTE: requires ggplot 2.2.0 
library("lubridate")

# monthly data series
library(zoo)
# extending ggplot2
library(scales)

#
library(cansim)

```

---

### data

**data source**



**Hours worked and labour productivity in the provinces and territories, 2017**

[The Daily, 2019-02-22](https://www150.statcan.gc.ca/n1/daily-quotidien/190222/dq190222e-eng.htm)


Labour productivity and related measures by business sector industry and by non-commercial activity consistent with the industry accounts

*  Table: 36-10-0480-01 (formerly CANSIM  383-0033)

*  https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1810020501



```{r}

data_36_10_0480_01 <- get_cansim("36-10-0480-01")

# -- this seemed unnecessary until I used year as a plotting variable
thedata <- data_36_10_0480_01 %>%
  mutate(REF_DATE = as_factor(REF_DATE))
#  mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) %>%
#  mutate(REF_DATE = year(REF_DATE))


thedata


```


#### data wrangling


geography list

* note: need stringr to allow filtering of `GeoUID`by 2-digit code (e.g., BC = 59)

```{r}

unique(thedata$GEO)

thedata %>%
#  filter(GeoUID == ) %>%
  group_by(GEO, GeoUID) %>%
  tally()

thedata %>%
  group_by(`Labour productivity and related measures`) %>%
  tally()

```





```{r}

# filter to have BC and labour productivity
thedata_BC <- thedata %>%
  filter(GEO == "British Columbia",
         `Labour productivity and related measures` == "Labour productivity")

thedata_BC


# filter to have two digit industry code & 1997

thedata_BC <- 
  thedata_BC %>%
  filter(str_length(`Classification Code for Industry`) == 6) %>%
  filter(REF_DATE == 1997)

thedata_BC


```

--- 

### plot

Uses a Cleveland dot plot
[https://uc-r.github.io/cleveland-dot-plots]()



1997 only

```{r}
# PLOT!
# basic
ggplot(thedata_BC, aes(x = VALUE, y = `Industry`)) + 
  geom_point()
#

# sorted by VALUE
ggplot(thedata_BC, aes(x = VALUE, y = reorder(`Industry`, VALUE))) + 
  geom_point()


```



Plot, 1997 and 2017

```{r}
# filter to have two digit industry code & 1997 + 2017

thedata_BC <- thedata %>%
  filter(
    str_length(`Classification Code for Industry`) == 6,
    REF_DATE %in% c(1997, 2017),
    GEO == "British Columbia",
    `Labour productivity and related measures` == "Labour productivity")

thedata_BC

```


```{r}

dataplot <- 
ggplot(thedata_BC, aes(x = VALUE, y = reorder(`Industry`, VALUE))) + 
  # plot points by year (REF_DATE), embiggen a bit
  geom_point(aes(color = REF_DATE), size = 3) +
  # add line to join points
  geom_line(aes(group = Industry))

dataplot

```




with more formatting

```{r}

dataplot2 <- dataplot +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_blank(),
    legend.text = element_text(size=10),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=9)
  )
#
dataplot2 

```

final version

```{r}

# experiments with ggplot2's new subtitle and caption options

cht_title <- as.character("Labour Productivity, B.C.")
cht_subtitle <- as.character("2012 chained dollars per hour")
cht_caption <- as.character("Source: Statistics Canada, CANSIM table 36-10-0480-01")

# add titles / X-Y axis labels
dataplot2 +
  ggtitle(cht_title, subtitle = cht_subtitle)

# add titles / X-Y axis labels
dataplot2 +
  labs(title = cht_title, subtitle = cht_subtitle)

# add titles / X-Y axis labels

LabourProd_plot <-
dataplot2 +
  labs(title = cht_title,
       subtitle = cht_subtitle,
       caption = cht_caption,
       x = "2012 chained dollars per hour", y = NULL)


LabourProd_plot

ggsave(filename = "LabourProd_plot.png", plot = LabourProd_plot,
       width = 8, height = 6)

```



