---
title: "Women in the workforce"
author: "Martin Monkman"
date: "2019-02-26"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

# more utilities
library(here)
library(janitor)


# plot theme
LFStheme <- 
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )




```



## Labour force participation: Women in the workforce

Request from media (unknown) about labour force participation of women age 20-44.


**Data source**

* Labour force labour_force_characteristics by sex and detailed age group, monthly, unadjusted for seasonality (x 1,000)

  - Table: 14-10-0017-01 (formerly CANSIM Table 282-0001)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410001701

* Labour force labour_force_characteristics by sex and detailed age group, annual (x 1,000)

  - Table: 14-10-0018-01 (formerly CANSIM  282-0002)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410001801
  


[Statistics Canada Open License Agreement](http://www.statcan.gc.ca/eng/reference/licence)





```{r include=FALSE}

# read the data - monthly
data_14_10_0017_01 <- get_cansim("14-10-0017-01")

data_14_10_0017_01 <- data_14_10_0017_01 %>%
    mutate(ref_date = ymd(ref_date, truncated = 2)) 

# read the data - annual
data_14_10_0018_01 <- get_cansim("14-10-0018-01")

#data_14_10_0018_01 <- data_14_10_0018_01 %>%
#    mutate(ref_date = ymd(ref_date, truncated = 2)) 


thedata <- data_14_10_0018_01 %>%
  clean_names()

unique(thedata$labour_force_characteristics)

thedata <-
thedata %>%
  mutate(geo = as_factor(geo),
         sex = as_factor(sex),
         labour_force_characteristics = as_factor(labour_force_characteristics),
         age_group = as_factor(age_group),
         ref_date = as.integer(ref_date)
  )


thedata


```


***

### Participation rate by age and sex


#### B.C. and Canada, Age == 25-44, sex = Female



```{r include=FALSE}

#
startdate <- 1976
prov.list <- c("British Columbia", "Canada")

# filter to have Canada
thedata_long <- thedata %>%
#  filter(ref_date >= startdate) %>%
#  filter(geo == "Canada",
  filter(geo %in% prov.list,
         age_group == "25 to 44 years",
         sex == "Females",
         labour_force_characteristics == "Participation rate")

### basic
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, group=geo)) + 
  geom_line() 

dataplot

dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=geo)) + 
  geom_line() 

dataplot



```


apply formatting
```{r}

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(50, 100), breaks = seq(50, 100, 10)) +
  LFStheme

dataplot2

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  scale_x_continuous(breaks = c(1976, seq(1980, 2015, by = 5), 2020)) +
#  theme(panel.grid.minor.x = c(1985, 1995, 2005, 2015)) +
  labs(title = "Labour force participation, females age 25-44, British Columbia & Canada",
       subtitle = "1976 - 2018 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 14-10-0018-01)", 
       x = NULL, y = "percentage")

dataplot2

# save the plot as a png file
#png("plot_labforpart_BCCan_female_25-44.png", width=1024, height=768, res=120)
#plot(dataplot2)
#dev.off()


```


For B.C. in `r max(thedata$ref_date)`, it was (pull value). Nationally, it was (pull value).

The participation rate rose rapidly in the mid-1970s through to about 1990, since then it has been rising more slowly. When Statistics Canada started the  Labour Force Survey in 1976, nationally the labour force participation rate for women aged 25-44 was 54%; it gained 10 percentage points over the next five years to reach 65.2% in 1981, and another 10 percentage points by 1988 (75.3%). The rate of increase since then has been much more modest.

```{r echo=FALSE}
thedata_long

dataplot2

```



***

#### B.C. and Canada, Age == 20-44, sex = Female and Male

```{r include=FALSE}

#
startdate <- 1976
prov.list <- c("British Columbia", "Canada")
sex.list <- c("Females", "Males")

# filters
thedata_long <- 
  thedata %>%
  filter(geo %in% prov.list,
         age_group == "25 to 44 years",
         sex %in% sex.list,
         labour_force_characteristics == "Participation rate")

thedata_long

### basic plot
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value)) + 
  geom_point()

dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, group=sex)) + 
  geom_line() 

dataplot

### basic plot
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=sex)) + 
  geom_line(size=1.5) +
  facet_grid(. ~ geo)

dataplot

```


```{r}

# with formatting applied

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(50, 100), breaks = seq(50, 100, 10)) +
  LFStheme


# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Labour force participation, age 25-44 by sex, British Columbia & Canada",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0002)", 
       x = NULL, y = "percentage")

dataplot2

# save the plot as a png file
png("plot_labforpart_BCCan_sex_25-44.png", width=1024, height=768, res=120)
plot(dataplot2)
dev.off()


```


The labour force participation of people aged 25-44 varies by sex. Males have higher participation rates, although the gap has closed significantly and continues to narrow.


```{r echo=FALSE}

thedata_long


dataplot2

```


***

#### B.C., Quebec, and Alberta

```{r include=FALSE}

#
startdate <- 1976
prov.list <- c("British Columbia", "Quebec", "Alberta")

# filters
thedata_long <- thedata %>%
  filter(geo %in% prov.list,
         age_group == "25 to 44 years",
         sex == "Females",
         labour_force_characteristics == "Participation rate")

### basic plot
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=geo)) + 
  geom_line(size=1.5) 

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(45, 100), breaks = seq(50, 100, 10)) +
  LFStheme

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Labour force participation, females age 25-44",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0002)", 
       x = NULL, y = "percentage")

# save the plot as a png file
png("plot_labforpart_BCprovs_female_25-44.png", width=1024, height=768, res=120)
plot(dataplot2)
dev.off()



thedata_long


dataplot2

```



***


## Labour force participation of women with children

### Background

On 2016-11-18, [Dr. Tammy Schirle](https://legacy.wlu.ca/homepage.php?grp_id=1805&f_id=31) posted [a series of tweets](https://twitter.com/tammyschirle/status/799616596750831616) via her twitter account (@tammyschirle), showing the proportion of Canadian women with young children, employed and at work, from 1976-2015.


**Data source**

Source: Statistics Canada, Labour Force Survey, Labour force characteristics by family age composition, annual

Table: 14-10-0120-01 (formerly CANSIM 282-0211)

DOI:   https://doi.org/10.25318/1410012001-eng

[Statistics Canada Open License Agreement](http://www.statcan.gc.ca/eng/reference/licence)




#### references

Kourtney Kobel and Tammy Schirle, ["The differential impact of universal child benefits on the labor supply of married and single mothers"](http://www.lcerpa.org/public/papers/LCERPA_2015_11.pdf)

Tyler Cowen, 2016, ["Why is female labor force participation down?"](http://marginalrevolution.com/marginalrevolution/2016/11/female-labor-force-participation.html) - U.S. measures



***



### Women with children in the workforce, Canada and British Columbia


```{r include=FALSE}

LFS_family <- get_cansim("14-10-0120-01")

LFS_family <- LFS_family %>%
  clean_names()

ls(LFS_family)

LFS_family <- LFS_family %>%
  mutate(geo = as_factor(geo),
         labour_force_characteristics = as_factor(labour_force_characteristics),
         family_age_composition  = as_factor(family_age_composition ),
         family_type  = as_factor(family_type ),
         ref_date = as.integer(ref_date)
  )


levels(LFS_family$labour_force_characteristics)
levels(LFS_family$family_age_composition)
levels(LFS_family$family_type)

```



```{r include=FALSE}


thedata <- LFS_family

#
startdate <- 1976
prov.list <- c("British Columbia", "Canada")


# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(geo %in% prov.list, 
         family_type == "Families with female reference person/spouse",
         family_age_composition == "Total families",
         labour_force_characteristics == "Participation rate")

### basic
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=geo)) + 
  geom_line(size=1.5) 

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  LFStheme

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, Canada & British Columbia",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

```

```{r echo=FALSE}

thedata_long

dataplot2

```


***


#### Canada, by age of child


```{r include=FALSE}

#
startdate <- 1976

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

# filters 
thedata_long <- thedata %>%
  filter(ref_date >= startdate) %>%
  filter(geo == "Canada", 
         family_type == "Families with female reference person/spouse",
         family_age_composition %in% family.comp.list,
         labour_force_characteristics == "Participation rate")

# PLOT!
# basic
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=family_age_composition)) + 
  geom_line(size=1.5) 

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Youngest child less than 3 years",
                         "Youngest child 3 to 5 years",
                         "Youngest child 6 to 15 years"),
                      values=c("#325A80", "#CCB550", "#000000")) +
  LFStheme

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, Canada",
       subtitle = "1976 - 2018 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

dataplot2

```



The labour force participation of women with children has increased in a similar way. In 1976 in British Columbia, fewer than one in three women (28%) whose youngest child was under the age of three participated in the workforce; by 2017 it was just over 71%. The labour force participation of women with older children has also increased: three-quarters of women whose youngest child is between three and five years of age (75%) and four out of five women whose youngest child is between 6 and 15 (83%).


```{r echo=FALSE}

thedata_long

dataplot2

```



---


#### British Columbia, by age of child


```{r include=FALSE}

#
startdate <- 1976

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

# filter to have BC
thedata_long <- thedata %>%
  filter(ref_date >= startdate) %>%
  filter(geo == "British Columbia", 
         family_type == "Families with female reference person/spouse",
         family_age_composition %in% family.comp.list,
         labour_force_characteristics == "Participation rate")

# PLOT!
# basic
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=family_age_composition)) + 
  geom_line(size=1.5) 


dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Youngest child less than 3 years",
                         "Youngest child 3 to 5 years",
                         "Youngest child 6 to 15 years"),
                      values=c("#325A80", "#CCB550", "#000000")) +
  LFStheme

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, British Columbia",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

```



```{r echo=FALSE}

thedata_long

dataplot2

```


---


#### British Columbia and Quebec, by age of child


```{r include=FALSE}

#
startdate <- 1976

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

thedata$family_age_composition <- ordered(thedata$family_age_composition, 
                              levels = c("Youngest child less than 3 years",
                                         "Youngest child 3 to 5 years",
                                         "Youngest child 6 to 15 years"))

prov.list <- c("British Columbia", "Quebec")

# filters 
thedata_long <- thedata %>%
  filter(ref_date >= startdate) %>%
  filter(geo %in% prov.list, 
         family_type == "Families with female reference person/spouse",
         family_age_composition %in% family.comp.list,
         labour_force_characteristics == "Participation rate")

# FACET PLOT!
#
# horizontal
dataplot <- ggplot(thedata_long, aes(x=ref_date, y=value, colour=geo)) + 
  geom_line(size=1.5) +
  facet_grid(. ~ family_age_composition)

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("British Columbia",
                         "Quebec"),
                      values=c("#325A80", "#CCB550")) +
  LFStheme +
  theme(
    legend.position=("bottom"), 
    legend.justification=c(1,0)
  )
#

# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, British Columbia & Quebec",
       subtitle = "1976 - 2018 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")


# save the plot as a png file
#png("plot_labforpart_BCQue_women-X-child-age.png", width=1024, height=768, res=120)
#plot(dataplot2)
#dev.off()

dataplot2

```



When Statistics Canada started measuring the national labour force in 1976, the participation rates of women with children was higher in B.C. than in Quebec, most notably for women whose youngest child was between 6 and 15 years of age. This persisted until the late 1990s, at which point the participation rate of women with children was higher in Quebec. This switch coincided with the introduction of policy measures regarding subsidized low-cost daycare.


```{r echo=FALSE}

thedata_long

dataplot2

```



Next steps: add some animation, as per Leonard Kiefer's US Labor Force Participation Rate chart:

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Labor force participation rate <a href="https://twitter.com/hashtag/dataviz?src=hash&amp;ref_src=twsrc%5Etfw">#dataviz</a> made with <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/gganimate?src=hash&amp;ref_src=twsrc%5Etfw">#gganimate</a> <a href="https://t.co/uSICoLjbIf">pic.twitter.com/uSICoLjbIf</a></p>&mdash; Leonard Kiefer (@lenkiefer) <a href="https://twitter.com/lenkiefer/status/1091352582776410113?ref_src=twsrc%5Etfw">February 1, 2019</a></blockquote>



-30-
