---
title: "LFS-partrate-by-age"
author: "Martin Monkman"
date: "November 21, 2016"
output: html_notebook:
  toc: true
  toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Seniors in the workforce

It has been reported that the labour force participation of older Canadians has been rising ... is this the case?

**Data source**

Source: Statistics Canada, Labour Force Survey. Reproduced and distributed on an "as is" basis with the permission of Statistics Canada.

CANSIM Table 282-0001
[Labour force survey estimates (LFS), by sex and detailed age group, unadjusted for seasonality: monthly]
(http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=2820001&pattern=labour+force+survey&tabMode=dataTable&srchLan=-1&p1=1&p2=-1) (downloaded 2018-03-23)


CANSIM Table 282-0002
[Labour force survey estimates (LFS), by sex and detailed age group, unadjusted for seasonality: annual](http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=2820002&pattern=labour+force+survey&tabMode=dataTable&srchLan=-1&p1=1&p2=-1) (downloaded 2018-03-23)


[Statistics Canada Open License Agreement](http://www.statcan.gc.ca/eng/reference/licence)


### Background

"The employment rate for seniors has more than doubled since 1988, from 6.7 per cent to 13.2 per cent." [Seniors and the generation spending gap](http://www.macleans.ca/society/life/seniors-and-the-generation-spending-gap/)


```{r}

# LFS
#
# install necessary packages
# utilities
library(readr)
library(zoo)
# plotting
library("ggplot2") # NOTE: requires ggplot 2.2.0 
library("scales")
# data manipulation
library("dplyr")
#
# set working directory
# setwd("J:/BCSTATS/INFOLINE/Infoline_2016-11-10_NHPI")
#
# read the data
# -- file 03270046-eng.csv download of CANSIM table 282-0002
thedata <- read_csv("../data/02820002-eng.csv")

```


## Age == 65+

```{r}

#
startdate <- 1976
prov.list <- c("Canada")

ls.str(thedata)

unique(thedata$CHARACTERISTICS)
unique(thedata$AGEGROUP)
unique(thedata$SEX)

# filter to have Canada
thedata_long <- thedata %>%
#  filter(Ref_Date >= startdate) %>%
  filter(GEOGRAPHY %in% prov.list,
         AGEGROUP == "65 years and over",
#         AGEGROUP == "15 to 24 years",
         SEX == "Both sexes",
         CHARACTERISTICS == "Participation rate (percentage)")


thedata_long

```



**Plot: Canada, 65+**

```{r}
### basic
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, group=GEOGRAPHY)) + 
  geom_line(size=1.5) 
#
# with formatting applied

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 15), breaks = seq(0, 15, 5)) +
#  scale_colour_manual(name=NULL,
#                      breaks=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"),
#                      labels=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"), 
#                      values=c("#325A80", "#CCB550", "#000000")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 +
  labs(title = "Labour force participation, persons age 65+, Canada",
       subtitle = "1976 - 2015 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0002)", 
       x = NULL, y = "percentage")

```

---

## British Columbia

```{r}
#
startdate <- 1976
prov.list <- c("British Columbia")

ls.str(thedata)


# filter to have British Columbia
thedata_long <- thedata %>%
#  filter(Ref_Date >= startdate) %>%
  filter(GEOGRAPHY %in% prov.list, 
         AGEGROUP == "65 years and over",
#         AGEGROUP == "15 to 24 years",
         SEX == "Both sexes",
         CHARACTERISTICS == "Participation rate (percentage)")

thedata_long


```



**Plot: Canada, 65+**


```{r}
### basic
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, group=GEOGRAPHY)) + 
  geom_line(size=1.5, colour = "#234275") 

dataplot
#
# with formatting applied

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 15), breaks = seq(0, 15, 5)) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Labour force participation, persons age 65+, British Columbia",
       subtitle = "1976 - 2015 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0002)", 
       x = NULL, y = "percentage")

# save the plot as a png file
png("plot_labforpart_BC_65plus.png", width=1024, height=768, res=120)
plot(dataplot2)
dev.off()


```


***

## Participation rate by age and sex


### Age == 20-44, Sex = Female

```{r}

#
startdate <- 1976
prov.list <- c("British Columbia", "Canada")

ls.str(thedata)

unique(thedata$CHARACTERISTICS)
unique(thedata$AGEGROUP)
unique(thedata$SEX)

# filter to have Canada
thedata_long <- thedata %>%
#  filter(Ref_Date >= startdate) %>%
  filter(GEOGRAPHY %in% prov.list,
         AGEGROUP == "25 to 44 years",
#         AGEGROUP == "15 to 24 years",
         SEX == "Females",
         CHARACTERISTICS == "Participation rate (percentage)")


thedata_long

```



```{r}
### basic
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=GEOGRAPHY)) + 
  geom_line(size=1.5) 
#
# with formatting applied

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(50, 100), breaks = seq(50, 100, 10)) +
#  scale_colour_manual(name=NULL,
#                      breaks=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"),
#                      labels=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"), 
#                      values=c("#325A80", "#CCB550", "#000000")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 +
  labs(title = "Labour force participation, females age 25-44, British Columbia & Canada",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0002)", 
       x = NULL, y = "percentage")

```
