---
title: "LFS-family-composition"
author: "Martin Monkman"
date: "November 19, 2016"
output: html_notebook:
    toc: true
    toc_float: true

editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Women with children in the workforce

### Background

On 2016-11-18, [Dr. Tammy Schirle](https://legacy.wlu.ca/homepage.php?grp_id=1805&f_id=31) posted [a series of tweets](https://twitter.com/tammyschirle/status/799616596750831616) via her twitter account (@tammyschirle), showing the proportion of Canadian women with young children, employed and at work, from 1976-2015.


**Data source**

Source: Statistics Canada, Labour Force Survey, Labour force survey estimates (LFS), by family type and family age composition ([CANSIM Table 282-0211](http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=2820211&&pattern=&stByVal=1&p1=1&p2=-1&tabMode=dataTable&csid=), downloaded 2017-03-26). Reproduced and distributed on an "as is" basis with the permission of Statistics Canada.

[Statistics Canada Open License Agreement](http://www.statcan.gc.ca/eng/reference/licence)

```{r}

# LFS
#
# install necessary packages
# utilities
library(readr)
library(zoo)
# plotting
library("ggplot2") # NOTE: requires ggplot 2.2.0 
library("scales")
# data manipulation
library("dplyr")
#
# set working directory
# setwd("J:/BCSTATS/INFOLINE/Infoline_2016-11-10_NHPI")
#
# read the data
# -- file 03270046-eng.csv download of CANSIM table 282-0211
#    "Labour force survey estimates (LFS), by family type and family age composition annual"                                             
# -- http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=2820211&&pattern=&stByVal=1&p1=1&p2=-1&tabMode=dataTable&csid=
thedata <- read_csv("../data/02820211-eng.csv")

thedata <- thedata %>%
  mutate(GEO = as.factor(GEO),
         FAMILYCOMP = as.factor(FAMILYCOMP),
         FAMILYTYPE = as.factor(FAMILYTYPE))

```


## Women with children in the workforce, Canada and British Columbia

```{r}

#
startdate <- 1976
prov.list <- c("British Columbia", "Canada")

ls.str(thedata)

unique(thedata$CHARACTERISTICS)
unique(thedata$FAMILYTYPE)
unique(thedata$FAMILYCOMP)

# filter to have BC and Canada
thedata_long <- thedata %>%
#  filter(Ref_Date >= startdate) %>%
  filter(GEO %in% prov.list, 
#  filter(GEO == "Canada", 
         FAMILYTYPE == "Families with female reference person/spouse",
         FAMILYCOMP == "Total families",
         CHARACTERISTICS == "Participation rate (percentage)")

thedata_long

```



**Plot: Canada and British Columbia**

```{r}
### basic
ggplot(thedata_long, aes(x=Ref_Date, y=Value, group=GEO)) + 
  geom_line()
#
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=GEO)) + 
  geom_line(size=1.5) 
dataplot


dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
#  scale_colour_manual(name=NULL,
#                      breaks=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"),
#                      labels=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"), 
#                      values=c("#325A80", "#CCB550", "#000000")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 +
  labs(title = "Women with children in the workforce, Canada & British Columbia",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

```


---


## Canada, by age of child


```{r}

#
startdate <- 1976

ls.str(thedata)

#unique(thedata$CHARACTERISTICS)
#unique(thedata$FAMILYTYPE)
unique(thedata$FAMILYCOMP)

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(Ref_Date >= startdate) %>%
#  filter(GEO %in% prov.list), 
  filter(GEO == "Canada", 
         FAMILYTYPE == "Families with female reference person/spouse",
         FAMILYCOMP %in% family.comp.list,
         CHARACTERISTICS == "Participation rate (percentage)")

thedata_long

```


```{r}
# PLOT!
# basic
ggplot(thedata_long, aes(x=Ref_Date, y=Value, group=FAMILYCOMP)) + 
  geom_line()
#
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=FAMILYCOMP)) + 
  geom_line(size=1.5) 
dataplot


dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Youngest child less than 3 years",
                         "Youngest child 3 to 5 years",
                         "Youngest child 6 to 15 years"),
#                      labels=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"), 
                      values=c("#325A80", "#CCB550", "#000000")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 +
  labs(title = "Women with children in the workforce, Canada",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

```



---


## British Columbia, by age of child


```{r}

#
startdate <- 1976

ls.str(thedata)

unique(thedata$CHARACTERISTICS)
unique(thedata$FAMILYTYPE)
unique(thedata$FAMILYCOMP)

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(Ref_Date >= startdate) %>%
#  filter(GEO %in% prov.list), 
  filter(GEO == "British Columbia", 
         FAMILYTYPE == "Families with female reference person/spouse",
         FAMILYCOMP %in% family.comp.list,
         CHARACTERISTICS == "Participation rate (percentage)")

thedata_long

```


```{r}
# PLOT!
# basic
ggplot(thedata_long, aes(x=Ref_Date, y=Value, group=FAMILYCOMP)) + 
  geom_line()
#
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=FAMILYCOMP)) + 
  geom_line(size=1.5) 
dataplot


dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Youngest child less than 3 years",
                         "Youngest child 3 to 5 years",
                         "Youngest child 6 to 15 years"),
#                      labels=c("Youngest child less than 3 years",
#                         "Youngest child 3 -5 years",
#                         "Youngest child 6 - 15 years"), 
                      values=c("#325A80", "#CCB550", "#000000")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, British Columbia",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

dataplot2

```



---


## British Columbia and Quebec, by age of child


```{r}

#
startdate <- 1976

ls.str(thedata)

unique(thedata$CHARACTERISTICS)
unique(thedata$FAMILYTYPE)
unique(thedata$FAMILYCOMP)

family.comp.list <- c("Youngest child less than 3 years",
                      "Youngest child 3 to 5 years",
                      "Youngest child 6 to 15 years")

thedata$FAMILYCOMP <- ordered(thedata$FAMILYCOMP, 
                              levels = c("Youngest child less than 3 years",
                                         "Youngest child 3 to 5 years",
                                         "Youngest child 6 to 15 years"))

prov.list <- c("British Columbia", "Quebec")

# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(Ref_Date >= startdate) %>%
#  filter(GEO %in% prov.list), 
  filter(GEO %in% prov.list, 
         FAMILYTYPE == "Families with female reference person/spouse",
         FAMILYCOMP %in% family.comp.list,
         CHARACTERISTICS == "Participation rate (percentage)")

thedata_long

```


```{r}
# FACET PLOT!
# vertical
ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=GEO)) + 
  geom_line(size=1.5) +
  facet_grid(FAMILYCOMP ~ .)
#
# horizontal
dataplot <- ggplot(thedata_long, aes(x=Ref_Date, y=Value, colour=GEO)) + 
  geom_line(size=1.5) +
  facet_grid(. ~ FAMILYCOMP)
dataplot

dataplot2 <- dataplot +
  scale_y_continuous(labels = comma, limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_colour_manual(name=NULL,
                      breaks=c("British Columbia",
                         "Quebec"),
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
#    legend.position=c(1,0), 
    legend.position=("bottom"), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 


# add titles / X-Y axis labels / caption
dataplot2 <- dataplot2 +
  labs(title = "Women with children in the workforce, British Columbia & Quebec",
       subtitle = "1976 - 2017 (annual)",
       caption = "Source: Statistics Canada (Labour Force Survey, CANSIM table 282-0211)", 
       x = NULL, y = "percentage")

dataplot2



# save the plot as a png file
png("plot_labforpart_BCQue_women-X-child-age.png", width=1024, height=768, res=120)
plot(dataplot2)
dev.off()


```




```{r}



```





---

## REFERENCES

Kourtney Kobel and Tammy Schirle, ["The differential impact of universal child benefits on the labor supply of married and single mothers"](http://www.lcerpa.org/public/papers/LCERPA_2015_11.pdf)

Tyler Cowen, 2016, ["Why is female labor force participation down?"](http://marginalrevolution.com/marginalrevolution/2016/11/female-labor-force-participation.html) - U.S. measures




