---
title: "Unemployment in British Columbia"
subtitle: "historic trend"
output: html_notebook
---


```{r setup}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)

library(cansim)
```

## LFS data by CMA

Note: CMA-level data starts in 2001

Labour force characteristics by census metropolitan area, annual 1
Frequency: Annual

Table: 14-10-0096-01 (formerly CANSIM 282-0129)

Geography: Census metropolitan area, Census metropolitan area part



Labour force characteristics by census metropolitan area, three-month moving average, seasonally adjusted (x 1,000)
Frequency: Monthly

Table: 14-10-0294-02

Geography: Canada, Province or territory, Census metropolitan area, Census metropolitan area part

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410029402



```{r}

df_1410029402 <- cansim::get_cansim("14-10-0294-02")

df_1410029402 <-
df_1410029402 %>% 
  janitor::clean_names()

unique(df_1410029402$geo)

bc_list <- c("British Columbia", "Victoria, British Columbia")

# filter BC employment at 2 digit NOC
df_lfs_bc <- df_1410029402 %>% 
  filter(geo %in% bc_list,
         labour_force_characteristics == "Unemployment")

```


## LFS, by province



Labour force characteristics by province, monthly, seasonally adjusted1 2 3
Frequency: Monthly

Table: 14-10-0287-03

Geography: Canada, Province or territory



Labour force characteristics by province, monthly, seasonally adjusted 1 2 3
Frequency: Monthly

Table: 14-10-0287-03

Geography: Canada, Province or territory



```{r}

df_1410028703 <- cansim::get_cansim("14-10-0287-03")

df_1410028703 <-
df_1410028703 %>% 
  janitor::clean_names()

unique(df_1410028703$geo)
unique(df_1410028703$labour_force_characteristics)


# filter BC employment at 2 digit NOC
df_lfs_bc <- df_1410028703 %>% 
  filter(geo == "British Columbia",
         labour_force_characteristics == "Unemployment rate")

```

single vector

```{r}

df_bc_unem_sa <- cansim::get_cansim_vector("v2064705", "1976-01-01")

df_bc_unem_sa <- janitor::clean_names(df_bc_unem_sa)

df_bc_unem_sa <- df_bc_unem_sa %>% 
  mutate(ref_date = ymd(ref_date))  


```


```{r}

p <- ggplot(df_bc_unem_sa, aes(x = ref_date, y = value)) +
  geom_line() +
  labs(title = "Unemployment rate: British Columbia",
       subtitle = "Seasonally adjusted",
       caption = "Source: Statistics Canada, Table 14-10-0287-03 v2064705",
       x = element_blank(),
       y = "percent unemployed"
  ) +
  ylim(0, 16)

p +
  annotate("rect", 
           xmin = ymd("1982-07-01"), xmax = ymd("1986-04-01"),
           ymin = 0, ymax = 16,
           fill = "blue", alpha = 0.1) +
  annotate("text", label = "44 consecutive months of unemployment of 13% or higher",
           x = ymd("1990-01-01"),
           y = 2.5,
           colour = "blue", alpha = 0.5,
           hjust = 0.0,
           size = 3) +
  annotate("segment",
           xend = ymd("1986-06-01"), x = ymd("1989-10-01"),
           y = 2.5, yend = 2.5,
           colour = "blue", size = 1.25, alpha = 0.5,
           arrow=arrow(), lineend = "round", linejoin = "mitre")

ggsave("BC_unemployment_1976-to-present.png")

```


Start 2000

```{r}

df_bc_unem_sa %>% 
  filter(ref_date >= "2000-01-01") %>% 
  ggplot(aes(x = ref_date, y = value)) +
  geom_line() +
  labs(title = "Unemployment rate: British Columbia",
       subtitle = "Seasonally adjusted",
       caption = "Source: Statistics Canada, Table 14-10-0287-03 v2064705",
       x = element_blank(),
       y = "percent unemployed"
  ) +
  ylim(0, 16)

ggsave("BC_unemployment_2000-to-present.png")


```

