---
title: "NHPI"
output: html_notebook
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


This set of scripts creates summary tables and graphs plotting the New Housing Price Index (NHPI) data collected and reported by Statistics Canada.

#### packages

```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library("ggplot2") # NOTE: requires ggplot 2.2.0 
#library("dplyr")

# monthly data series
library(zoo)
# extending ggplot2
library("scales")


```

---

### data


ARCHIVED DATA SERIES TO 2016-12 
read the data
- file 03270046-eng.csv download of CANSIM table 327-0046
- http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=3270046&&pattern=&stByVal=1&p1=1&p2=1&tabMode=dataTable&csid=

#thedata <- read_csv("data/03270046-eng.csv")

#### download and save

NEW DATA SERIES
Table 327-0056
New housing price index
monthly (index, 201612=100)
- http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=3270056&&pattern=&stByVal=1&p1=1&p2=-1&tabMode=dataTable&csid= 

```{r}

thedata <- read_csv("../data/03270056-eng.csv")

```


#### data wrangling

```{r}


# fix dates
class(thedata$Ref_Date)
thedata$DATE <- as.Date(as.yearmon(thedata$Ref_Date, format="%Y/%m"))
#

startdate <- as.Date("2007-01-01")


# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(DATE >= startdate) %>%
  filter(GEO == c("British Columbia", "Canada"), INDEX == "Total (house and land)")

thedata_long


```

--- 

### plot


```{r}
# PLOT!
# basic
ggplot(thedata_long, aes(x=DATE, y=Value, group=GEO)) + 
  geom_line()
#
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=DATE, y=Value, colour=GEO)) + 
  geom_line(size=1.5) 
dataplot

```

with more formatting

```{r}

dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(80, 110)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Canada", "British Columbia"),
                      labels=c("Canada", "British Columbia"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 

```

final version

```{r}

# experiments with ggplot2's new subtitle and caption options

NHPI_title <- as.character("New Housing Price Index, Canada & B.C.")
NHPI_subtitle <- as.character("December 2016 = 100")
NHPI_caption <- as.character("Source: Statistics Canada, CANSIM table 327-0056")

# add titles / X-Y axis labels
dataplot2 +
  ggtitle(NHPI_title, subtitle = NHPI_subtitle)

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title, subtitle = NHPI_subtitle)

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       x = NULL, y = "NHPI")

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI") 

# final version
# add titles / X-Y axis labels / caption
NHPI_plot <- dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI (Dec. 2016 = 100)") 

NHPI_plot

ggsave(filename = "NHPI_plot.png", plot = NHPI_plot,
       width = 8, height = 6)

```



** DAVE WAS HERE **