---
title: "NHPI"
output:
  html_document:
    df_print: paged
---
<!--
Copyright 2018-2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


This set of scripts creates summary tables and graphs plotting the New Housing Price Index (NHPI) data collected and reported by Statistics Canada.

This version (updated 2018-12-13) uses the `cansim` package to pull the data from CANSIM.


#### packages



```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library("ggplot2") # NOTE: requires ggplot 2.2.0 

library("lubridate")

# monthly data series
library(zoo)
# extending ggplot2
library(scales)

#
library(cansim)

```

---

### data


ARCHIVED DATA SERIES TO 2016-12 
read the data
- file 03270046-eng.csv download of CANSIM table 327-0046
- http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=3270046&&pattern=&stByVal=1&p1=1&p2=1&tabMode=dataTable&csid=


```{r}
#thedata <- read_csv("data/03270046-eng.csv")

```


#### download and save

**data source**



[The Daily, 2018-12-13](https://www150.statcan.gc.ca/n1/daily-quotidien/181213/dq181213b-eng.htm)

NEW DATA SERIES

Table: 18-10-0205-01 (formerly CANSIM  327-0056)

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1810020501




```{r}

#thedata <- read_csv("../data/03270056-eng.csv")

table_id = "18-10-0205-01"


thedata <- get_cansim(table_id)
thedata <- thedata %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

get_cansim_table_overview(table_id)


thedata


```




#### data wrangling


geography list

* note: need stringr to allow filtering of `GeoUID`by 2-digit code (e.g., BC = 59)

```{r}

unique(thedata$GEO)

thedata %>%
#  filter(GeoUID == ) %>%
  group_by(GEO, GeoUID) %>%
  tally()

```





```{r}

startdate <- as.Date("2007-01-01")


# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(REF_DATE >= startdate) %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `New housing price indexes` == "Total (house and land)")

thedata_long


```

--- 

### plot


```{r}
# PLOT!
# basic
ggplot(thedata_long, aes(x=REF_DATE, y=VALUE, group=GEO)) + 
  geom_line()
#
# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=REF_DATE, y=VALUE, colour=GEO)) + 
  geom_line(size=1.5) 
dataplot

```

with more formatting

```{r}

dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(80, 120)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Canada", "British Columbia"),
                      labels=c("Canada", "British Columbia"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 

```

final version

```{r}

# experiments with ggplot2's new subtitle and caption options

NHPI_title <- as.character("New Housing Price Index, Canada & B.C.")
NHPI_subtitle <- as.character("December 2016 = 100")
NHPI_caption <- as.character("Source: Statistics Canada, CANSIM table 18-10-0205-01")

# add titles / X-Y axis labels
dataplot2 +
  ggtitle(NHPI_title, subtitle = NHPI_subtitle)

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title, subtitle = NHPI_subtitle)

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       x = NULL, y = "NHPI")

# add titles / X-Y axis labels
dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI") 

# final version
# add titles / X-Y axis labels / caption
NHPI_plot <- dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI (Dec. 2016 = 100)") 

NHPI_plot

ggsave(filename = "NHPI_plot.png", plot = NHPI_plot,
       width = 8, height = 6)

```


***

### Vancouver house and land version



```{r}

startdate <- as.Date("2007-01-01")


# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(REF_DATE >= startdate) %>%
  filter(GEO %in% c("Vancouver, British Columbia"), 
         `New housing price indexes` %in% c("House only", "Land only"))

thedata_long


```

the plot

```{r}

# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=REF_DATE, y=VALUE, 
                                     colour=`New housing price indexes`)) + 
  geom_line(size=1.5) 
dataplot


dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(80, 120)) +
  scale_colour_manual(name=NULL,
                      breaks=c("House only", "Land only"),
                      labels=c("House only", "Land only"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 




# set chart titles and labels

NHPI_title <- as.character("New Housing Price Index, Vancouver: house and land")
NHPI_subtitle <- as.character("December 2016 = 100")
NHPI_caption <- as.character("Source: Statistics Canada, CANSIM table 18-10-0205-01")

# final version
# add titles / X-Y axis labels / caption
NHPI_Vancouver_plot <- dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI (Dec. 2016 = 100)") 

NHPI_Vancouver_plot

ggsave(filename = "NHPI_Vancouver_plot.png", plot = NHPI_Vancouver_plot,
       width = 8, height = 6)


```


***

***

### Halifax house and land version



```{r}

startdate <- as.Date("2007-01-01")


# filter to have BC and Canada
thedata_long <- thedata %>%
  filter(REF_DATE >= startdate) %>%
  filter(GEO %in% c("Halifax, Nova Scotia"), 
         `New housing price indexes` %in% c("House only", "Land only"))

thedata_long


```

the plot

```{r}

# with formatting applied
dataplot <- ggplot(thedata_long, aes(x=REF_DATE, y=VALUE, 
                                     colour=`New housing price indexes`)) + 
  geom_line(size=1.5) 
dataplot


dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(70, 110)) +
  scale_colour_manual(name=NULL,
                      breaks=c("House only", "Land only"),
                      labels=c("House only", "Land only"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 




# set chart titles and labels

NHPI_title <- as.character("New Housing Price Index, Halifax: house and land")
NHPI_subtitle <- as.character("December 2016 = 100")
NHPI_caption <- as.character("Source: Statistics Canada, CANSIM table 18-10-0205-01")

# final version
# add titles / X-Y axis labels / caption
NHPI_Halifax_plot <- dataplot2 +
  labs(title = NHPI_title,
       subtitle = NHPI_subtitle,
       caption = NHPI_caption, 
       x = NULL, y = "NHPI (Dec. 2016 = 100)") 

NHPI_Halifax_plot

ggsave(filename = "NHPI_Halifax_plot.png", plot = NHPI_Halifax_plot,
       width = 8, height = 6)


```


***

# Vector number table

```{r}

thedata %>%
  group_by(VECTOR,
           GEO,
           `New housing price indexes`) %>%
  tally()


thedata %>%
  group_by(VECTOR,
           GEO,
           `New housing price indexes`) %>%
  filter(GEO %in% c("Canada", "British Columbia")) %>%
  tally()

```



