---
title: "components of population change"
output: html_document
---

```{r}
# packages
library(tidyverse)
#
library(cansim)
#
library(here)
here()

library(janitor)

```

objective: code to create two CSV files, with migration by regional district and development region

published to BC Data Catalogue here:

* https://catalogue.data.gov.bc.ca/dataset/migration-by-development-region-and-regional-district


reference:

* https://catalogue.data.gov.bc.ca/dataset/inter-provincial-and-international-migration

For international data, total number of individuals leaving BC = Emigrants + Net Temporary Emigrants – Returning Emigrants. 
Total number moving to BC = International Immigrants + Net Non Permanent Residents.

## Regional District

### source

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710014001

Components of population change by census division, 2016 boundaries1, 2, 3, 4, 5
Frequency: Annual

Table: 17-10-0140-01

Release date: 2022-01-13

Geography: Province or territory, Census division


```{r read_cansim}
## bogs down

#table_num <- "17-10-0140-01"
#  
#get_cansim_table_overview(table_num)
#
#df_components <- get_cansim(table_num)
#
```

18.4 million rows, 17 columns ... takes a moment

NOTE: turned off as R chunk ... to run, re-insert the ` ```{r} ` initiation

```

df_components <- read_csv(here("data", "17100140-eng", "17100140.csv"))

df_components <- df_components %>% 
  janitor::clean_names()

write_rds(df_components, here("data", "17100140-eng", "17100140.rds"))

df_components <- read_rds(here("data", "17100140-eng", "17100140.rds"))

ls.str(df_components)

### filter for BC total

df_components_bc <- df_components %>% 
  filter(geo == "British Columbia")

df_components_bc

write_rds(df_components_bc, here("data", "17100140-eng", "df_components_bc.rds"))

df_components_bc <- read_rds(here("data", "17100140-eng", "df_components_bc.rds"))


### filter for regional districts

df_components_bc_cd <- df_components %>% 
  filter(str_ends(geo, ", British Columbia")) %>% 
  mutate("SGC" = stringr::str_sub(dguid, -4, -1))


```

### read BC regional districts


Regional district number is the last 2 digits of the "dguid" variable -- last 4 digits include the "59" indicating BC. This appears elsewhere as the "SGC" variable.


```

# filter for BC regional districts

df_components_bc_cd <- df_components %>% 
  filter(str_ends(geo, ", British Columbia")) %>% 
  mutate(sgc = stringr::str_sub(dguid, -4, -1))
  

write_rds(df_components_bc_cd, here("data", "17100140-eng", "df_components_bc_cd.rds"))

```


```{r read_cd}

df_components_bc_cd <- read_rds(here("data", "17100140-eng", "df_components_bc_cd.rds"))

max(df_components_bc_cd$ref_date)

```


just for fun, filter for max year and Capital 

```{r}
df_components_bc_cd %>% 
  filter(ref_date == "2020/2021") %>% 
  filter(str_starts(geo, "Capital")) 

```

2742 rows! Yikes

```{r}

# what are the "components_of_population_growth"?

unique(df_components_bc_cd$components_of_population_growth)

```

drop births and deaths

```{r}

df_components_bc_cd_crd06 <- df_components_bc_cd %>% 
  filter(ref_date == "2006/2007") %>% 
  filter(sgc == 5917) %>%
  #
  filter(age_group == "All ages") %>% 
  filter(sex == "Both sexes") %>% 
  #
  filter(components_of_population_growth != "Residual deviation") %>% 
  filter(components_of_population_growth != "Births") %>% 
  filter(components_of_population_growth != "Deaths") 

df_components_bc_cd_crd06

```

Now pivot that sucker and see how it compares to the posted CSV file...

```{r}

df_components_bc_cd_crd06 %>% 
  select(ref_date, sgc, geo, components_of_population_growth, value) %>% 
  pivot_wider(names_from = components_of_population_growth,
              values_from = value)
```

## Development Region

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710013801

Components of population change by economic region, 2016 boundaries1, 2, 3, 4, 5
Frequency: Annual

Table: 17-10-0138-01

Release date: 2022-01-13

Geography: Province or territory, Economic region

