---
title: "SEPH"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Survey of Employment, Payrolls and Hours


This set of scripts creates summary tables and graphs plotting the monthly SEPH data collected and reported by Statistics Canada.

#### packages

```{r packages}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
#library(forcats)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)


### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)

```

---

## Background

The Daily, 2018-12-20

* https://www150.statcan.gc.ca/n1/daily-quotidien/181220/dq181220b-eng.htm

The Daily, 2018-11-29

* https://www150.statcan.gc.ca/n1/daily-quotidien/181129/dq181129c-eng.htm

The Daily, 2018-10-25

* https://www150.statcan.gc.ca/n1/daily-quotidien/181025/dq181025a-eng.htm


**Important:** "In general, changes in weekly earnings reflect a number of factors, including wage growth; changes in the composition of employment by industry, occupation and level of job experience; and average hours worked per week."



## -- Data

There are multiple (13) data tables associated with the SEPH:
https://www150.statcan.gc.ca/n1/daily-quotidien/181025/dq181025a-cansim-eng.htm

The principal tables used here are:

* Employment and average weekly earnings (including overtime) for all employees by province and territory, monthly, seasonally adjusted

  - Table: 14-10-0223-01 (formerly CANSIM  281-0063)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410022301


* Average weekly hours for employees paid by the hour, by industry, monthly, unadjusted for seasonality

  - Table: 14-10-0255-01 (formerly CANSIM  281-0032)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410025501
  


### Read the data


**NOTE:**

This script uses the [R package `cansim`](https://github.com/mountainMath/cansim)


```{r data_read}

data_14_10_0223_01 <- get_cansim("14-10-0223-01")

data_14_10_0223_01 <- data_14_10_0223_01 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 
  

```


#### understanding the data 

```{r}

filenamenum <- "data_14_10_0223_01"

print(glue("TABLE:  ", filenamenum))

ls.str(data_14_10_0223_01)

print("----")
print("values of Estimate")
unique(data_14_10_0223_01$Estimate)

print("----")
print("values of NAICS")
unique(data_14_10_0223_01$`North American Industry Classification System (NAICS)`)


print("----")
print("most recent date in series (variable `lastmonth`)")
lastmonth <- max(data_14_10_0223_01$REF_DATE)
lastmonth

```


```{r}
data_14_10_0223_01 %>%
  filter(GEO == "Canada",
         Estimate == "Employment for all employees") %>%
  filter(REF_DATE == lastmonth)

data_14_10_0223_01 %>%
  filter(GEO == "British Columbia",
         Estimate == "Employment for all employees") %>%
  filter(REF_DATE == lastmonth)


```


#### vector numbers

Industrial aggregate including unclassified businesses	- Canada: v79310773 / BC: v79311043

Goods producing industries	- Canada: v79310775 / BC: v79311045

Service producing industries	- Canada: v79310785 / BC: v79311055


```{r}


data_14_10_0223_01 %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `North American Industry Classification System (NAICS)` == 
           "Industrial aggregate excluding unclassified businesses",
         Estimate == "Employment for all employees") %>%
  group_by(GEO) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
  arrange(GEO, desc(REF_DATE)) %>%
  select(REF_DATE, GEO, Estimate,
         `North American Industry Classification System (NAICS)`,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  filter(REF_DATE >= "2008-01-01")


data_14_10_0223_01 %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `North American Industry Classification System (NAICS)` == 
           "Industrial aggregate excluding unclassified businesses", 
         Estimate == "Average weekly earnings including overtime for all employees") %>%
  group_by(GEO) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
  arrange(GEO, desc(REF_DATE)) %>%
  select(REF_DATE, GEO, Estimate,
         `North American Industry Classification System (NAICS)`,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  filter(REF_DATE >= "2008-01-01")

```


### Plot: BC vs Canada


```{r}

dataplot <- data_14_10_0223_01 %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         `North American Industry Classification System (NAICS)` == "Industrial aggregate excluding unclassified businesses", 
         Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(REF_DATE >= as.Date("2008-01-01")) %>%
  ggplot(aes(x=REF_DATE, y=VALUE, group = GEO, colour = GEO)) + 
    geom_line(size=1.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)
  
dataplot

```




```{r}

dataplot2 <- dataplot +
  ylim(700, 1050) +
  scale_y_continuous(labels = comma, limits = c(700, 1050)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Canada", "British Columbia"),
                      labels=c("Canada", "British Columbia"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=10)
  )
#
dataplot2 

# experiments with ggplot2's new subtitle and caption options

SEPH_earnings_title <- as.character("Average weekly earnings including overtime for all employees")
#SEPH_earnings_subtitle <- as.character("January 2010 = 100")
SEPH_earnings_caption <- as.character("Source: Statistics Canada, CANSIM table 14-10-0223-01")
SEPH_earnings_yaxis <- as.character("Dollars")

# add titles / X-Y axis labels / caption
SEPH_earnings_plot <- dataplot2 +
  labs(title = SEPH_earnings_title,
#       subtitle = SEPH_earnings_subtitle,
       caption = SEPH_earnings_caption, 
       x = NULL, y = SEPH_earnings_yaxis) 

SEPH_earnings_plot

ggsave(filename = "SEPH_earnings_plot.png", plot = SEPH_earnings_plot,
       width = 8, height = 6)

```




---



## -- Sectors 

Year-over-year change in average weekly earnings in the 10 largest sectors

https://www150.statcan.gc.ca/n1/daily-quotidien/181129/cg-c002-eng.htm


```{r}

# Make a B.C. table for the most recent month

BC_lastmonth <- data_14_10_0223_01 %>%
  filter(GEO == "British Columbia") %>%
  group_by(`North American Industry Classification System (NAICS)`,
           Estimate) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
#  arrange(desc(VALUE)) %>%
  select(REF_DATE, GEO, Estimate,
         `North American Industry Classification System (NAICS)`,
         VECTOR,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  filter(REF_DATE == lastmonth)


BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees")

BC_lastmonth %>%
  filter(Estimate == "Employment for all employees") 

BC_lastmonth %>%
  filter(Estimate == "Employment for all employees") %>%
  arrange(desc(VALUE))


# remove sector aggregations
# note use of %!in% defined earlier
BC_lastmonth %>%
  na.omit() %>%
  filter(Estimate == "Employment for all employees") %>%
  arrange(desc(VALUE)) %>%
  filter(VECTOR %!in% 
           c("v79311043", "v79311046", "v79311055",
             "v79311054", "v79311045")
           ) 
```




#### Plot: top 10 sectors

Retail trade	 6
Professional, scientific and technical 13
Accommodation and food services 8
ALL SECTORS
Wholesale trade 16
Public admin 12
Admin and support 15
Construction 10
Health care and social assistance	7
Educational services 9
Manufacturing 11
bonus: Transportation and warehousing 14


```{r}

  
top10list <- c("Retail trade", "Health care and social assistance",
               "Accommodation and food services", "Educational services",
               "Construction", "Manufacturing",
               "Public administration", 
               "Professional, scientific and technical services",
               "Transportation and warehousing",
               "Administrative and support, waste management and remediation services")

top10list_fac <- as_factor(top10list)


BC_lastmonth <- 
BC_lastmonth %>%
  mutate(NAICS_desc =
           case_when(
             `North American Industry Classification System (NAICS)` == 
               "Health care and social assistance" ~ 
               "Health care & social assistance",
             `North American Industry Classification System (NAICS)` == 
               "Accommodation and food services" ~ 
               "Accommodation & food services",
             `North American Industry Classification System (NAICS)` == 
               "Professional, scientific and technical services" ~ 
               "Professional, scientific & technical services",
             `North American Industry Classification System (NAICS)` == 
               "Transportation and warehousing" ~ 
               "Transportation & warehousing",
             `North American Industry Classification System (NAICS)` == 
               "Administrative and support, waste management and remediation services" ~ 
               "Administrative & support",
             TRUE ~ `North American Industry Classification System (NAICS)`
           ))




BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  arrange(desc(MOM_pct)) %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list)


BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  arrange(desc(YOY_pct)) %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list)

```




Month over month plot

```{r}

dataplot <- 
BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  arrange(desc(MOM_pct)) %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list) %>%
  ggplot(aes(x = VALUE,
             y = reorder(NAICS_desc, VALUE) 
             )) + 
    geom_point()

dataplot

dataplot <- 
BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list) %>%
  ggplot(aes(x = MOM_pct,
             y = reorder(NAICS_desc, MOM_pct))) +
    geom_point() +
    geom_segment(aes(yend=NAICS_desc), xend=0, colour="grey50")

dataplot

```




Year over year plot

```{r}


dataplot <- 
BC_lastmonth %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list) %>%
  ggplot(aes(x = YOY_pct,
             y = reorder(NAICS_desc, YOY_pct))) +
    geom_point(colour="grey50", size = 4) +
    geom_segment(aes(yend=NAICS_desc), xend=0, colour="grey50", size = 1.25)

dataplot

```


```{r}

dataplot2 <- dataplot +
  scale_x_continuous(labels = comma, 
                     limits = c(-4, 8),
                     breaks = seq(-4, 8, 2)) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=10)
  )
#
dataplot2 

```



```{r}


# experiments with ggplot2's new subtitle and caption options

SEPH_earnings_title <- as.character("B.C. Average weekly earnings\nincluding overtime")
SEPH_earnings_subtitle <- as.character("October 2018 compared to October 2017")
SEPH_earnings_caption <- as.character("Source: Statistics Canada, CANSIM table 14-10-0223-01")
SEPH_earnings_xaxis <- as.character("Year-over-year percent change")

# add titles / X-Y axis labels / caption
SEPH_earnings_plot <- dataplot2 +
  labs(title = SEPH_earnings_title,
       subtitle = SEPH_earnings_subtitle,
       caption = SEPH_earnings_caption, 
       y = NULL, x = SEPH_earnings_xaxis) 

SEPH_earnings_plot

ggsave(filename = "SEPH_earnings_change_plot.png", plot = SEPH_earnings_plot,
       width = 8, height = 6)

```


#### Plot: facet by sector

```{r}
# Make a B.C. table for the top ten sectors, going back to 2008-01

BC_decade <- data_14_10_0223_01 %>%
  filter(GEO == "British Columbia") %>%
  group_by(`North American Industry Classification System (NAICS)`,
           Estimate) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
#  arrange(desc(VALUE)) %>%
  select(REF_DATE, GEO, Estimate,
         `North American Industry Classification System (NAICS)`,
         VECTOR,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  filter(REF_DATE >= "2008-01-01")


BC_decade %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees")

BC_decade %>%
  filter(Estimate == "Employment for all employees") 

BC_decade %>%
  filter(Estimate == "Employment for all employees") %>%
  arrange(desc(VALUE))

```



```{r}

BC_decade %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list) %>%
  ggplot(aes(x = REF_DATE,
             y = VALUE, 
             colour = `North American Industry Classification System (NAICS)`)) +
    geom_line()

dataplot <-
BC_decade %>%
  filter(Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(`North American Industry Classification System (NAICS)` %in% top10list) %>%
  ggplot(aes(x = REF_DATE,
             y = VALUE)) +
    geom_line()
  

dataplot +
  facet_grid(. ~ `North American Industry Classification System (NAICS)`)

dataplot +
  facet_grid(cols = vars(`North American Industry Classification System (NAICS)`))

dataplot +
  facet_grid(rows = vars(`North American Industry Classification System (NAICS)`))



```



Sort the facets from smallest value to highest

see https://ggplot2.tidyverse.org/reference/facet_grid.html

https://forcats.tidyverse.org/reference/fct_reorder.html



-30-