<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->

---
title: "SEPH - construction"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

# Survey of Employment, Payrolls and Hours


This set of scripts creates summary tables and graphs plotting the monthly SEPH data collected and reported by Statistics Canada.

#### packages

```{r, packages}

# tidyverse
library(tidyverse)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library(scales)

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

# utilities
library(brickr)
library(here)
library(janitor)

### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)

```

---

## Background


The Daily, 2019-07-25: Payroll employment, earnings and hours, May 2019
https://www150.statcan.gc.ca/n1/daily-quotidien/190725/dq190725a-eng.htm

The Daily, 2019-01-30 (November 2018)
https://www150.statcan.gc.ca/n1/daily-quotidien/190130/dq190130a-eng.htm 

The Daily, 2018-12-20 (October 2018)

* https://www150.statcan.gc.ca/n1/daily-quotidien/181220/dq181220b-eng.htm

The Daily, 2018-11-29

* https://www150.statcan.gc.ca/n1/daily-quotidien/181129/dq181129c-eng.htm

The Daily, 2018-10-25

* https://www150.statcan.gc.ca/n1/daily-quotidien/181025/dq181025a-eng.htm


**Important:** "In general, changes in weekly earnings reflect a number of factors, including wage growth; changes in the composition of employment by industry, occupation and level of job experience; and average hours worked per week."



## -- Data

There are multiple (13) data tables associated with the SEPH:
https://www150.statcan.gc.ca/n1/daily-quotidien/181025/dq181025a-cansim-eng.htm

The principal tables used here are:

* Employment and average weekly earnings (including overtime) for all employees by province and territory, monthly, seasonally adjusted

  - Table: 14-10-0223-01 (formerly CANSIM  281-0063)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410022301


* Average weekly hours for employees paid by the hour, by industry, monthly, unadjusted for seasonality

  - Table: 14-10-0255-01 (formerly CANSIM  281-0032)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410025501
  


### Read the data


**NOTE:**

This script uses the [R package {cansim}](https://github.com/mountainMath/cansim)


```{r data_read}

#data_14_10_0223_01 <- get_cansim("14-10-0223-01", refresh = TRUE)
data_14_10_0223_01 <- get_cansim("14-10-0223-01")

```


```{r}

LFS_data <- data_14_10_0223_01 %>%
  janitor::clean_names() %>%
  mutate(ref_date = ymd(ref_date, truncated = 2)) %>%
  filter(geo == "British Columbia",
         estimate == "Employment for all employees",
         north_american_industry_classification_system_naics == "Construction [23]")

LFS_data <- LFS_data %>%
  filter(month(ref_date) == 5,
         year(ref_date) > 2009)

LFS_data

ls.str(LFS_data)

```

## plot

```{r}

LFS_plot <- ggplot(LFS_data) +
  aes(x=ref_date, y=value) + 
  geom_brick_col() 
#  scale_fill_brick() 
#  coord_brick() +
#  theme_brick()

LFS_plot

```

plot, with formatting

```{r}

LFS_plot +
  labs(title = "Employment in B.C.'s construction industry continues to grow",
       subtitle = "The month of May, 2010-2019",
       caption = "Statistics Canada.  Table  14-10-0201-01   Employment by industry, monthly, unadjusted for seasonality") +
  xlab(element_blank()) +
  ylab("number of persons") +
  scale_y_continuous(labels = comma)

```

```{r}

LFS_plot +
  labs(title = "Employment in B.C.'s construction industry continues to grow",
       subtitle = "The month of May, 2010-2019",
       caption = "Statistics Canada.  Table  14-10-0201-01   Employment by industry, monthly, unadjusted for seasonality") +
  xlab("year") +
  ylab("number of persons") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%Y",
               date_breaks = "1 year", 
               minor_breaks = NULL) 
#  theme(panel.grid.minor.x = element_blank())

```


```{r}

LFS_plot +
  labs(title = "Employment in B.C.'s construction industry continues to grow",
       subtitle = "The number of persons employed during the month of May, 2010-2019",
       caption = "Source: Statistics Canada. Table 14-10-0201-01; Employment by industry, monthly, unadjusted for seasonality") +
  xlab("year") +
  ylab("number of persons") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%Y",
               date_breaks = "1 year", 
               minor_breaks = NULL) 
               

ggsave(filename = "SEPH_construction_plot.png", 
       width = 8, height = 6)


```

-30-