---
title: "SEPH_medianwage_animate"
author: "Martin Monkman"
date: "December 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggthemes)
library(ggrepel)

# Package to pull in cansim data, courtesy of Jens von Bergman and Dmitry Shkolnik
library(cansim) 

# Now update to the newest gganimate version
#install.packages("devtools")
#devtools::install_github('thomasp85/gganimate')
library(gganimate)

```

## source

This is plagarized from Blake Schaffer's ["Making animated charts with gganimate"](https://www.blakeshaffer.ca/post/making-animated-charts-with-gganimate/)


Schaffer created a chart showing the per capita GDP of each Canadian province, using Statistics Canada data from 1981-2017.

For this makeover, we'll look at median income values, drawn from the Survey of Employment, Payrolls and Hours (SEPH).


### gganimate reference material

[github page](https://github.com/thomasp85/gganimate)



```{r cars}
provlist <- c("British Columbia","Alberta","Saskatchewan","Manitoba","Ontario","Quebec","New Brunswick","Nova Scotia","Prince Edward Island","Newfoundland and Labrador")

short_provs <- function(df){
  df <- df %>%
    filter(GEO %in% c("British Columbia","Alberta","Saskatchewan","Manitoba","Ontario","Quebec","New Brunswick","Nova Scotia","Prince Edward Island","Newfoundland and Labrador"))
  df <- df %>%
    mutate(GEO.short = case_when(
      GEO=="British Columbia" ~ "BC",
      GEO=="Alberta" ~ "AB",
      GEO=="Saskatchewan" ~ "SK",
      GEO=="Manitoba" ~ "MB",
      GEO=="Ontario" ~ "ON",
      GEO=="Quebec" ~ "QC",
      GEO=="New Brunswick" ~ "NB",
      GEO=="Prince Edward Island" ~ "PE",
      GEO=="Nova Scotia" ~ "NS",
      GEO=="Newfoundland and Labrador" ~ "NL"
    ))
  df$GEO.short <- factor(df$GEO.short, levels=c("BC","AB","SK","MB","ON","QC","NB","PE","NS","NL"))
}

```


get the data


```{r data_read}

data_14_10_0223_01 <- get_cansim("14-10-0223-01")

data_14_10_0223_01 <- data_14_10_0223_01 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 


AWE <- data_14_10_0223_01 %>%
  filter(GEO %in% provlist, 
         `North American Industry Classification System (NAICS)` == "Industrial aggregate excluding unclassified businesses", 
         Estimate == "Average weekly earnings including overtime for all employees")  %>%
  mutate(year = year(REF_DATE),
         month = month(REF_DATE)) %>%
  select(GEO, REF_DATE, year, month, VALUE)

AWE$GEO.short <- short_provs(AWE)

AWE

AWE_annual <- AWE %>%
  group_by(GEO.short, year) %>%
  summarise(AWE_ann_avg = mean(VALUE)) %>%
  mutate(year = as.integer(year))

AWE_annual

```



plot B.C. trend line

```{r}

AWE %>%
  filter(GEO.short == "BC") %>%
  ggplot(aes(x = REF_DATE, y = VALUE)) +
  geom_line()

```


animated plot using annual average AWE

```{r}


p <- ggplot(AWE_annual, aes(GEO.short, AWE_ann_avg, fill = GEO.short, label=GEO.short)) +
  geom_col()+
#  geom_text(vjust=-.1)+
  scale_fill_viridis_d(name="")+
  scale_y_continuous(labels=scales::dollar)+
#  theme_tufte(base_size = 14)+
  theme(legend.position = "none") +
#        axis.text.x = element_blank(),
#        axis.ticks = element_blank())+
# Here comes the gganimate specific bits
  labs(title = "Year: {frame_time}", x = "", y = "Average weekly earnings") +
  transition_time(year) +
  ease_aes('linear')        

p

animate(p, nframes = 100, fps = 5, end_pause = 20) # use anim_save(filename) to save

anim_save("SEPH_AWE_anim_01.gif")

```



animated plot using monthly values

_note: this doesn't work very well -- the value is a sum of the months of the year, and the labs(title) is a decimalized fraction of the year


```{r}


ggplot(AWE, aes(GEO.short, VALUE, fill = GEO.short, label=GEO.short)) +
  geom_col()+
#  geom_text(vjust=-.1)+
  scale_fill_viridis_d(name="")+
  scale_y_continuous(labels=scales::dollar)+
#  theme_tufte(base_size = 14)+
  theme(legend.position = "none") +
#        axis.text.x = element_blank(),
#        axis.ticks = element_blank())+
# Here comes the gganimate specific bits
  labs(title = "Year: {frame_time}", x = "", y = "Average weekly earnings") +
  transition_time(year) +
  ease_aes('linear')        

```



2nd attempt at monthly values, but with transition_time(month)

```{r}


ggplot(AWE, aes(GEO.short, VALUE, fill = GEO.short, label=GEO.short)) +
  geom_col()+
#  geom_text(vjust=-.1)+
  scale_fill_viridis_d(name="")+
  scale_y_continuous(labels=scales::dollar)+
#  theme_tufte(base_size = 14)+
  theme(legend.position = "none") +
#        axis.text.x = element_blank(),
#        axis.ticks = element_blank())+
# Here comes the gganimate specific bits
  labs(title = "Year: {frame_time}", x = "", y = "Average weekly earnings") +
  transition_time(month) +
  ease_aes('linear')        

```


