<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->



---
title: "Service exports"
output: html_notebook
---

## Canada's International Transactions in Services

http://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=1536


#### CANSIM tables

List:

* https://www150.statcan.gc.ca/n1/daily-quotidien/181019/dq181019e-cansim-eng.htm 

* all Canada only (i.e. no provincial breakdown)

International transactions in services, travel by category and geographical area, annual (x 1,000,000)

* Table: 36-10-0004-01 (formerly CANSIM  376-0031)

International transactions in services, transportation by category, annual (x 1,000,000)

* Table: 36-10-0005-01 (formerly CANSIM  376-0032)

International transactions in services, commercial services by category, annual (x 1,000,000)

* Table: 36-10-0006-01 (formerly CANSIM  376-0033)

International transactions in services, by selected countries, annual (x 1,000,000)

* Table: 36-10-0007-01 (formerly CANSIM  376-0036)

International transactions in services, commercial services, by North American Industry Classification System (NAICS), annual (x 1,000,000)

* Table: 36-10-0022-01 (formerly CANSIM  376-0109) 




#### The Daily_

Canada's international trade in services, 2017

* 2017-10-19

* https://www150.statcan.gc.ca/n1/daily-quotidien/181019/dq181019e-eng.htm 


Canada's international trade in services, 2016

* 2017-10-12

* https://www150.statcan.gc.ca/n1/daily-quotidien/171012/dq171012c-eng.htm


Denis Caron (613-808-2278; denis.caron@canada.ca), International Accounts and Trade Division.


##### Methods & additional reports

https://www150.statcan.gc.ca/n1/en/catalogue/13-605-X201800154966

https://www150.statcan.gc.ca/n1/en/catalogue/13-605-X201800154965 




### data extract


```{r packages}

library(cansim)

library(dplyr)
library(tidyr)
library(forcats) # for factor variables
library(readr)

```

International transactions in services, commercial services, by North American Industry Classification System (NAICS)

```{r data_read}

# read the data 
DF_source <- get_cansim("36-10-0022-01")

DF_source

```


examine the dataframe

```{r}

unique(DF_source$GEO)

print("-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-")

unique(DF_source$`North American Industry Classification System (NAICS)`)

print("-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-")

unique(DF_source$`Commercial services by category`)



```

