---
title: "Survey of Manufacturing"
author: "Martin Monkman"
date: "September 18, 2018"
output: html_document
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This set of scripts creates summary tables and graphs plotting the Monthly Survey of Manufacturing data collected and reported by Statistics Canada.

#### packages

```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)

# monthly data series
#library(zoo)
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

```

---

The Daily, 2018-09-18

* https://www150.statcan.gc.ca/n1/daily-quotidien/180918/dq180918a-eng.htm 


### data

National
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1610004701

Provincial (sales only)
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1610004801


Read the data

* note: two files, one for Canada and another for the provinces and territories

```{r}

mfg_CAN <- get_cansim("16-10-0047-01")
mfg_prov <- get_cansim("16-10-0048-01")


```


#### data wrangling

filter `mfg_CAN` to

* include only total manufacturing & seasonally adjusted


filter `mfg_prov` to 

* include only British Columbia

  - then only total manufacturing & seasonally adjusted



```{r}

# Canada

ls.str(mfg_CAN)

mfg_CAN_totalSA <- mfg_CAN %>%
  mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) %>%
  filter(`North American Industry Classification System (NAICS)` == "Manufacturing",
         `Seasonal adjustment` == "Seasonally adjusted",
         `Principal statistics` == "Sales of goods manufactured (shipments)") %>%
  select(REF_DATE, 
         GEO,
         `North American Industry Classification System (NAICS)`,
         `Classification Code for North American Industry Classification System (NAICS)`,
         `Seasonal adjustment`,
         `Principal statistics`,
         VALUE
         )


# provinces

ls.str(mfg_prov)

unique(mfg_prov$GEO)

mfg_BC <- mfg_prov %>%
  mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) %>%
  filter(GEO == "British Columbia")

mfg_BC_totalSA <- mfg_BC %>%
  filter(`North American Industry Classification System (NAICS)` == "Manufacturing",
         `Seasonal adjustment` == "Seasonally adjusted") %>%
  select(REF_DATE, 
         GEO,
         `North American Industry Classification System (NAICS)`,
         `Classification Code for North American Industry Classification System (NAICS)`,
         `Seasonal adjustment`,
         `Principal statistics`,
         VALUE
         )
  


```



Add index values

* indexpct 

```{r}

indexrefdate <- "2010-01-01"

# Canada

index100_CAN <- as.integer(
  mfg_CAN_totalSA %>%
  filter(REF_DATE == indexrefdate) %>%
  select(VALUE)
)

mfg_CAN_totalSA <- mfg_CAN_totalSA %>%
  mutate(indexpct = (VALUE / index100_CAN) * 100)


# BC

index100_BC <- as.integer(
  mfg_BC_totalSA %>%
  filter(REF_DATE == "2010-01-01") %>%
  select(VALUE)
)

mfg_BC_totalSA <- mfg_BC_totalSA %>%
  mutate(indexpct = (VALUE / index100_BC) * 100)

```



#### Plot!



join tables


```{r}

mfg_CANBC_totalSA <- mfg_CAN_totalSA %>%
  full_join(mfg_BC_totalSA)


```



plot with both Canada and B.C.



```{r}

mfg_CANBC_totalSA %>%
  filter(REF_DATE >= "2010-01-01") %>%
  ggplot(aes(x=REF_DATE, y=indexpct, colour=GEO)) + 
  geom_line()

```

* plus format




```{r}

dataplot <- 
  mfg_CANBC_totalSA %>%
  filter(REF_DATE >= "2010-01-01") %>%
  ggplot(aes(x=REF_DATE, y=indexpct, colour=GEO)) + 
    geom_line(size=1.5) 
dataplot

```

with more formatting

```{r}

dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(100, 175)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Canada", "British Columbia"),
                      labels=c("Canada", "British Columbia"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=12),
    axis.text = element_text(size=12)
  )
#
dataplot2 

```

final version

```{r}

# experiments with ggplot2's new subtitle and caption options

Mfg_title <- as.character("Manufacturing Sales (seasonally adjusted), Canada & B.C.")
Mfg_subtitle <- as.character("January 2010 = 100")
Mfg_caption <- as.character("Source: Statistics Canada, CANSIM tables 16-10-0047 & 16-10-0048")

# add titles / X-Y axis labels / caption
Mfg_plot <- dataplot2 +
  labs(title = Mfg_title,
       subtitle = Mfg_subtitle,
       caption = Mfg_caption, 
       x = NULL, y = "Index") 

Mfg_plot

ggsave(filename = "Mfg_plot.png", plot = Mfg_plot,
       width = 8, height = 6)

```



### Canadian detail

Looking at the past three months, sorted by total growth. 

(**Note:** this uses the seasonally adjusted data!)

```{r}


mfg_CAN_industry <- 
  mfg_CAN %>%
  mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) %>%
  filter(str_detect(
    !!(mfg_CAN$`Classification Code for North American Industry Classification System (NAICS)`), 
        "\\[...]")) %>%
  filter(`Seasonal adjustment` == "Seasonally adjusted",
         `Principal statistics` == "Sales of goods manufactured (shipments)") %>%
  group_by(`North American Industry Classification System (NAICS)`) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  select(REF_DATE, 
         `North American Industry Classification System (NAICS)`, 
         `Classification Code for North American Industry Classification System (NAICS)`,
         VALUE, MOM_val, MOM_pct) %>%
  arrange(REF_DATE, desc(MOM_val)) 


mfg_CAN_industry %>%
  filter(REF_DATE == "2018-07-01")

mfg_CAN_industry %>%
  filter(REF_DATE == "2018-06-01")

mfg_CAN_industry %>%
  filter(REF_DATE == "2018-05-01")


```

