---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Wholesale trade


This set of scripts creates summary tables and graphs plotting the monthly Wholesale Trade data collected and reported by Statistics Canada.

#### packages

```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

```

---

## Background

The Daily, 2018-09-24

* https://www150.statcan.gc.ca/n1/daily-quotidien/180924/dq180924a-eng.htm



## data

There are 3 data tables associated with this:
https://www150.statcan.gc.ca/n1/daily-quotidien/180924/dq180924a-cansim-eng.htm



* Wholesale sales, price and volume, by industry, seasonally adjusted (x 1,000,000)
  
  - national data only
  
  - Table: 20-10-0003-01 (formerly CANSIM  081-0018)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2010000301

* Wholesale trade, sales (x 1,000)

  - contains provincial series, both unadjusted and adjusted, with industry details
  
  - Table: 20-10-0074-01 (formerly CANSIM  081-0011)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2010007401
  
* Wholesale trade, inventories (x 1,000)

  - national data only

  - Table: 20-10-0076-01 (formerly CANSIM 081-0012)

  - https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=2010007601
  


###

Read the data


```{r}

data_20_10_0074 <- get_cansim("20-10-0074-01")

data_20_10_0074 <- data_20_10_0074 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 
  

```



try some filtering

```{r}

ls.str(data_20_10_0074)


data_20_10_0074 %>%
  filter(GEO %in% c("Canada", "British Columbia"),
    `Adjustments` == "Seasonally adjusted",
    `Classification Code for North American Industry Classification System (NAICS)` == "[41]") %>%
  group_by(GEO, `Adjustments`) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
  arrange(GEO, desc(REF_DATE)) %>%
  select(REF_DATE, GEO,
         Adjustments,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  ungroup() %>%
#  filter(REF_DATE >= "2017-07-01", 
  filter(REF_DATE %in% c(as.Date("2018-07-01"), as.Date("2018-06-01"), as.Date("2017-07-01")), 
         `Adjustments` == "Seasonally adjusted")


data_20_10_0074 %>%
  filter(`Adjustments` == "Seasonally adjusted",
    `Classification Code for North American Industry Classification System (NAICS)` == "[41]") %>%
  group_by(GEO, `Adjustments`) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
  select(REF_DATE, GEO,
         Adjustments,
         VALUE, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  ungroup() %>%
  filter(REF_DATE >= "2018-07-01") %>%
  arrange(desc(MOM_pct))


```




```{r}


BC_data_0074 <- 
 data_20_10_0074 %>%
  filter(GEO == "British Columbia",
    `Adjustments` %in% c("Unadjusted", "Seasonally adjusted") ) %>%
  group_by(GEO, `Adjustments`,
           `Classification Code for North American Industry Classification System (NAICS)`) %>%
  mutate(MOM_val = lag(VALUE),
         MOM_pct = ((VALUE / lag(VALUE, n = 1)) - 1) * 100,
         MOM_chg = (VALUE - lag(VALUE, n = 1)) ) %>%
  mutate(YOY_val = lag(VALUE, n = 12),
         YOY_pct = ((VALUE / lag(VALUE, n = 12)) - 1) * 100,
         YOY_chg = (VALUE - lag(VALUE, n = 12)) ) %>%
  arrange(VECTOR, REF_DATE) %>%
  ungroup()

BC_data_0074

BC_data_0074 %>%
  filter(REF_DATE == "2018-07-01")

write_csv(BC_data_0074, "BC_data_0074.csv")

```




***

old and plagiarized code, uncertain functionality

```{r}


BC_data_0074 %>%
  filter(str_detect(
    !!(BC_data_0074$`Classification Code for North American Industry Classification System (NAICS)`), 
        "\\[...]")) 

BC_data_0074 %>%
  filter(str_detect(
    !!(BC_data_0074$`Classification Code for North American Industry Classification System (NAICS)`), 
        "\\[...]")) %>%
  filter(REF_DATE == "2018-07-01") %>%
  select(REF_DATE, GEO,
         `North American Industry Classification System (NAICS)`,
         `Adjustments`,
         VALUE,
         MOM_pct, MOM_val, MOM_chg)



```


plot: BC trend

```{r}

data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  select(REF_DATE, VALUE, MOM_pct, MOM_val) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2006-01-01")


data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5) 


```


```{r}

unique(year(data_10_0011$REF_DATE))

data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)

```


```{r}

dataplot <- data_10_0011 %>%
  filter(GEO %in% c("British Columbia"),
    `Beneficiary detail` == "Regular benefits",
    Sex == "Both sexes",
    `Age group` == "15 years and over") %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#filter(REF_DATE == "2018-07-01")
  filter(REF_DATE >= "2005-01-01") %>%
  ggplot(aes(x=REF_DATE, y=VALUE)) + 
    geom_line(size=1.5, colour="#325A80") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)
  



dataplot2 <- dataplot +
  #  ylim(3500000, 6500000) +
  scale_y_continuous(labels = comma, limits = c(30000, 100000)) +
#  scale_colour_manual(name=NULL,
#                      breaks=c("Canada", "British Columbia"),
#                      labels=c("Canada", "British Columbia"), 
#                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=10)
  )
#
dataplot2 

# experiments with ggplot2's new subtitle and caption options

EI_title <- as.character("Employment Insurance beneficiaries (seasonally adjusted), B.C.")
EI_subtitle <- as.character("January 2010 = 100")
EI_caption <- as.character("Source: Statistics Canada, CANSIM table 14-10-0011-01")
EI_yaxis <- as.character("Number of EI beneficiaries")

# add titles / X-Y axis labels / caption
EI_plot <- dataplot2 +
  labs(title = EI_title,
#       subtitle = EI_subtitle,
       caption = EI_caption, 
       x = NULL, y = EI_yaxis) 

EI_plot

ggsave(filename = "EI_plot.png", plot = EI_plot,
       width = 8, height = 6)

```


### misc stray code from previous

```{r}
NAICScodes <- unique(BC_data_0074$`Classification Code for North American Industry Classification System (NAICS)`)
str_detect(NAICScodes, "\\[...]")

BC_data_0008 %>%
    filter(str_detect(!!NAICScodes, "\\[...]")) 



BC_data_0008 %>%
  filter(REF_DATE == "2017-07-01")

BC_data_0008 %>%
  filter(REF_DATE == "2018-07-01")

```



```{r}


unique(data_10_0013$GeoUID)

CMA_CA_list <- c("[11124]",
                 "[59]", "[59A]", "[932]", "[915]", "[933]", "[935]", 
                 "[59B]", "[944]", "[930]", "[943]", "[905]", 
                 "[975]", "[937]", "[977]", "[925]", "[938]", "[939]", 
                 "[913]", "[940]", "[945]", "[970]", "[955]", "[952]", 
                 "[920]", "[934]", "[965]", "[918]", "[950]", 
                 "[59C]", "[59996]", "[59997]", "[59998]", "[59999]")


filenamenum <- "data_10_0013"

print(glue("TABLE:  ", filenamenum))

ls.str(data_10_0013)



data_10_0013 %>%
  filter(GeoUID %in% CMA_CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
#  arrange(desc(REF_DATE)) %>%
  filter(REF_DATE == "2018-07-01")



data_10_0013 %>%
  filter(GeoUID %in% CMA_CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_val)) %>%
  filter(REF_DATE == "2018-07-01")

CMA_list <- c("[59A]", "[932]", "[915]", "[933]", "[935]")

data_10_0013 %>%
  filter(GeoUID %in% CMA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_pct)) %>%
  filter(REF_DATE == "2018-07-01")


CA_list <- c("[59B]", "[944]", "[930]", "[943]", "[905]", 
             "[975]", "[937]", "[977]", "[925]", "[938]", "[939]", 
             "[913]", "[940]", "[945]", "[970]", "[955]", "[952]", 
             "[920]", "[934]", "[965]", "[918]", "[950]")


data_10_0013 %>%
  filter(GeoUID %in% CA_list) %>%
  group_by(GEO) %>%
  mutate(MOM_pct = VALUE / lag(VALUE),
         MOM_val = VALUE - lag(VALUE)) %>%
  arrange(desc(MOM_pct)) %>%
  filter(REF_DATE == "2018-07-01")

```


