---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


# Wholesale trade


This set of scripts creates summary tables and graphs plotting the monthly Wholesale Trade data collected and reported by Statistics Canada.

#### packages

```{r}

# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)

# utilities
library(here)
library(janitor)
library(brickr)

```

---

## Background

The Daily, 2018-09-24

* https://www150.statcan.gc.ca/n1/daily-quotidien/180924/dq180924a-eng.htm


The Daily, 2019-08-22

 * https://www150.statcan.gc.ca/n1/daily-quotidien/190822/dq190822b-eng.htm


## The data

There are 3 data tables associated with this:
https://www150.statcan.gc.ca/n1/daily-quotidien/180924/dq180924a-cansim-eng.htm



* Wholesale sales, price and volume, by industry, seasonally adjusted (x 1,000,000)
  
  - national data only
  
  - Table: 20-10-0003-01 (formerly CANSIM  081-0018)
  
  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2010000301

* Wholesale trade, sales (x 1,000)

  - contains provincial series, both unadjusted and adjusted, with industry details
  
  - Table: 20-10-0074-01 (formerly CANSIM  081-0011)

  - https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2010007401
  
* Wholesale trade, inventories (x 1,000)

  - national data only

  - Table: 20-10-0076-01 (formerly CANSIM 081-0012)

  - https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=2010007601
  


### Read the data


```{r}
# not working 2019-08-22
#data_20_10_0074 <- get_cansim("20-10-0074-01")
#data_20_10_0074 <- get_cansim("081-0011")

data_20_10_0074 <- read_csv(here("data", "20100074.csv"))


data_20_10_0074 <- data_20_10_0074 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2))  %>%
    janitor::clean_names() 

  

```



try some filtering
```{r}
ls.str(data_20_10_0074)

```



```{r}



data_20_10_0074 %>%
  filter(geo %in% c("Canada", "British Columbia"),
    adjustments == "Seasonally adjusted",
    north_american_industry_classification_system_naics == "Wholesale trade [41]") %>%
  group_by(geo, adjustments) %>%
  mutate(MOM_val = lag(value),
         MOM_pct = ((value / lag(value, n = 1)) - 1) * 100,
         MOM_chg = (value - lag(value, n = 1)) ) %>%
  mutate(YOY_val = lag(value, n = 12),
         YOY_pct = ((value / lag(value, n = 12)) - 1) * 100,
         YOY_chg = (value - lag(value, n = 12)) ) %>%
  arrange(geo, desc(ref_date)) %>%
  select(ref_date, geo,
         adjustments,
         value, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  ungroup() %>%
#  filter(REF_DATE >= "2017-07-01", 
  filter(ref_date %in% c(as.Date("2019-09-01"), as.Date("2019-05-01"), as.Date("2018-06-01")), 
         adjustments == "Seasonally adjusted")


data_20_10_0074 %>%
  filter(
    adjustments == "Seasonally adjusted",
    north_american_industry_classification_system_naics == "Wholesale trade [41]") %>%
  group_by(geo, adjustments) %>%
  mutate(MOM_val = lag(value),
         MOM_pct = ((value / lag(value, n = 1)) - 1) * 100,
         MOM_chg = (value - lag(value, n = 1)) ) %>%
  mutate(YOY_val = lag(value, n = 12),
         YOY_pct = ((value / lag(value, n = 12)) - 1) * 100,
         YOY_chg = (value - lag(value, n = 12)) ) %>%
  select(ref_date, geo,
         adjustments,
         value, 
         MOM_val, MOM_pct, MOM_chg,
         YOY_val, YOY_pct, YOY_chg) %>%
  ungroup() %>%
  filter(ref_date >= "2019-06-01") %>%
  arrange(desc(MOM_pct))


```

### B.C. data


```{r}


BC_data_0074 <- 
 data_20_10_0074 %>%
  filter(geo == "British Columbia",
    adjustments %in% c("Unadjusted", "Seasonally adjusted") ) %>%
  group_by(geo, adjustments,
           north_american_industry_classification_system_naics) %>%
  mutate(MOM_val = lag(value),
         MOM_pct = ((value / lag(value, n = 1)) - 1) * 100,
         MOM_chg = (value - lag(value, n = 1)) ) %>%
  mutate(YOY_val = lag(value, n = 12),
         YOY_pct = ((value / lag(value, n = 12)) - 1) * 100,
         YOY_chg = (value - lag(value, n = 12)) ) %>%
  arrange(vector, ref_date) %>%
  ungroup()

ls.str(BC_data_0074)

BC_data_0074 %>%
  filter(ref_date == "2019-06-01")

write_csv(BC_data_0074, "BC_data_0074_2019-08-22.csv")

```



```{r}

library(stringr)

str_extract("Farm product merchant wholesalers [411]", "\\[[0-9]+\\]$")
str_extract("Textile, clothing and footwear merchant wholesalers [4141]", "\\[[0-9]+\\]$")

head(
  str_extract(BC_data_0074$north_american_industry_classification_system_naics, "\\[[0-9]+\\]$")
)

###

str_remove("Farm product merchant wholesalers [411]", " merchant wholesalers \\[[0-9]+\\]$")
str_remove("Textile, clothing and footwear merchant wholesalers [4141]", " merchant wholesalers \\[[0-9]+\\]$")

###
```


```{r}

BC_data_0074 <- BC_data_0074 %>%
  mutate(naics_code = 
           str_extract(north_american_industry_classification_system_naics,
                       "\\[[0-9]+\\]$"),
         naics_short = 
           str_remove(north_american_industry_classification_system_naics, 
                      " merchant wholesalers \\[[0-9]+\\]$")
  )

# filter 3 digit NAICS industries

BC_data_0074 %>%
  filter(str_detect(naics_code, "\\[[0-9]{3}\\]$")) %>%
  filter(ref_date == "2019-06-01",
         adjustments == "Unadjusted") %>%
  arrange(desc(value))

```


### plot: BC trend

```{r}

BC_data_0074 %>%
  filter(str_detect(naics_code, "\\[[0-9]{3}\\]$")) %>%
  filter(ref_date == "2019-06-01",
         adjustments == "Unadjusted") %>%
  arrange(desc(value)) %>%
  #
  ggplot() +
  aes(x=as_factor(naics_code), y=value) + 
  geom_brick_col() +
  scale_fill_brick() +
  coord_brick() +
  theme_brick()

```

```{r}

BC_data_0074 %>%
  filter(str_detect(naics_code, "\\[[0-9]{3}\\]$")) %>%
  filter(ref_date == "2019-06-01",
         adjustments == "Unadjusted") %>%
  arrange(value) %>%
  #
  ggplot() +
  aes(x=as_factor(north_american_industry_classification_system_naics), y=value) + 
  geom_brick_col() +
  scale_fill_brick() +
  coord_brick_flip() +
  theme_brick() +
  theme(legend.position = "none")

```



```{r}





BC_data_0074 %>%
  filter(str_detect(naics_code, "\\[[0-9]{3}\\]$")) %>%
  filter(ref_date == "2019-06-01",
         adjustments == "Unadjusted") %>%
  arrange(value) %>%
  #
  ggplot() +
  aes(x=as_factor(naics_short), y=value) + 
  geom_brick_col() +
  scale_fill_brick() +
  coord_brick_flip() +
  theme_brick() +
  theme(legend.position = "none") +
  xlab(element_blank())


```



